---
title: "GeneralityRpf"
author: "Venus Kuo"
date: "April 2, 2017"
output: html_document
---
# Based on growthcurve_code.R Written by: M. Larsen (2013/07/18) and M. Muscarella (2015/11/24) # 


## 1) Questions
Is Rpf effects general on taxa unrelated to Micrococcus luteus?

## 2) Methods

Microbial Growth Curves on KBS enviromental strains

1. 701(-) Bacteriodetes, Sphingobacteriaceae, Pedobacter Gram -
2. 702 (+),Actinobacteria, Micrococcaceae, Arthrobacter Gram +
3. 703 (+),Actinobacteria, Micrococcaceae, Arthrobacter Gram +
4. 705 (-), Proteobacteria, Rhodospirillaceae, Azospirillium Gram - KmR. Tc and Gm inhibited growth
5. 706 (+),Actinobact., Mycobacteriaceae, Mycobacterium Gram +
6. 710 (-), Proteobacteria, Pseudomonadaceae, Pseudomonas Gram - AmR, StR, SpR, RfR
7. 711 (-), Proteobacteria, Oxalobacteriaceae, Janthinobacterium Gram -AmR, SpR
8. 712 (-), Proteobacteria, Comamonadaceae, Variovorax Gram -AmR, KmR, StR, GmR, SpR, RfR
9. 714 (+), Actinobacteria, Micrococcaceae, Micrococcus Gram +
10. 715 (+), Actinobacteria, Microbacteriaceae, Curtobacterium Gram +
11. 724 (+), Actinobacteria, Nocardiaceae, Rhodococcus Gram +
12. 812 (+). Bacillus subtilis SSB2 WT Gram +


# Set working directory #

```{r, echo=FALSE}
# Set working directory #
rm(list = ls())
getwd()
setwd("C:/Users/Venus/Github/Rpf/generality/")

# Require or install packages #
package.list <- c('vegan', 'ggplot2', 'gplots', 'coin', 'vegan', 'nlme' ,'data.table', 'plyr', 'reshape', 'reshape2', 'ggplot2') 
for (package in package.list) {
  if (!require(package, character.only=T, quietly=T)) { 
    install.packages(package)
    library(package, character.only=T)
} }
```
# Extract umax, L, and A from Growth_Curve code # 

```{r, echo = FALSE}



```

## 3) Description of the data set 



# Load data set #


```{r}
# Output of my raw data run saved as "ModifiedGompEquationOutput.csv" #
data <- read.csv("GrowthCurveData.csv", header = T)

# Subset from data # 
umax <- subset(data, select = c("Genus", "Treatment", "umax"))
A <- subset(data, select = c("Genus", "Treatment", "A"))
L <- subset(data, select = c("Genus", "Treatment", "L"))

# Mean values of umax, A, and L # 
umax.mean <- ddply(umax, c("Genus", "Treatment"), summarise, mean=mean(umax))
A.mean <- ddply(A, c("Genus", "Treatment"), summarise, mean=mean(A))
L.mean <- ddply(L, c("Genus", "Treatment"), summarise, mean=mean(L))
```

## Data Visualization ##  

```{r}
# Bar graph showing the difference in umax between treatments for each strain # 


# Plot the umax difference between treatments for each strain # 


```

## Data Anaylsis ##

```{r}


```


