---
title: "Rpf: Growth Curve"
author: "Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

Termination of dormancy in soil bacteria with resuscitation promoting factor (Rpf)

## Clear workspace and set directory

```{r setup}
rm(list=ls())
getwd()
#setwd("~/GitHub/Rpf")
knitr::opts_knit$set(root.dir = '~/../GitHub/Rpf') 
```

## Load packages and functions

```{r}
require("png")
require("grid")
require("plyr")
require("dplyr")

sem <- function(x) sqrt(var(x)/length(x))
cv <- function(x) sd(x)/mean(x)
```

## Load growth data

This included monitoring of dormant KBS0714 populations in Klett flasks. 
Measured OD600 daily on flasks with (478 nM) vs. without (protein bufffer) Rpf.

```{r}
growth <- read.table("data/growth.curve/growth.curve.txt", sep="\t", header=TRUE)
```

## Make figure

```{r}
png(filename="~/GitHub/Rpf/figures/Fig3.png",
    width = 1200, height = 900, res = 96*2) 

plot.new()
par(mar = c(7, 7, 5, 7))

# Add +Rpf tubes
plot(growth[,1], growth[,6], xlim = c(-25, 625), ylim = c(-0.5, 4), type = "l", 
     lty = 2, col = "black", lwd = 2, ylab = "", xlab = "", 
     cex.lab = 1.5, las = 1, yaxt = "n", xaxt = "n")
box(lwd = 2)

points(growth[,1], growth[,7], type = "l", lty = 2, col = "black", lwd = 2)
points(growth[,1], growth[,8], type = "l", lty = 2, col = "black", lwd = 2)
points(growth[,1], growth[,9], type = "l", lty = 2, col = "black", lwd = 2)

# Add -Rpf tubes
points(growth[,1], growth[,2], type = "l", lty = 3, col = "black", lwd = 2)
points(growth[,1], growth[,3], type = "l", lty = 3, col = "black", lwd = 2)
points(growth[,1], growth[,4], type = "l", lty = 3, col = "black", lwd = 2)
points(growth[,1], growth[,5], type = "l", lty = 3, col = "black", lwd = 2)

# Add axes
axis(side = 2, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
   labels = c("0", "1", "2", "3", "4"), 
        at = c(0, 1, 2, 3, 4))

axis(side = 4, labels = F, lwd.ticks = 2, 
   at = c(0, 1, 2, 3, 4))

axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1, mgp = c(1, 1, 0),
    labels = c(0, 200, 400, 600), at = c(0, 200, 400, 600))

axis(side = 3, labels = F, lwd.ticks = 2, las = 1, cex.axis = 1, 
   at = c(0, 200, 400, 600))

# Add axis labels
mtext('Biomass (OD600)', side = 2, outer = TRUE, cex = 1.5, 
      line = -4.5, adj = 0.6)

mtext('Time (h)', side = 1, outer = TRUE, cex = 1.5, 
      line = -4, adj = 0.5)

text(400, 3.6, labels = "+Rpf", cex = 1)
text(540, 3.6, labels = "-Rpf", cex = 1)

# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("./figures/Fig3.png")

grid.raster(img)
```

## Analyze growth curve parameters

# Load parameters derived from modified Gompertz model

```{r}
g.parms <- read.delim("data/growth.curve/parameters.txt", sep=",", header=TRUE)
```

# Compare summary statistics of growth curve parameters

```{r}
Treatment <- c("Rpf-", "Rpf-","Rpf-","Rpf-", "Rpf+","Rpf+","Rpf+","Rpf+")
g.parms$Treatment <- Treatment

# Maximum growth rate (umax)
umax.init <- ddply(g.parms, c("Treatment"), summarise,
                   mean=mean(umax), sem=sd(umax)/sqrt(length(umax)))

umax <- transform(umax.init, lower=mean-sem, upper=mean+sem)

# Lag time
lag.init <- ddply(g.parms, c("Treatment"), summarise,
                   mean=mean(L), sem=sd(L)/sqrt(length(L)))
lag <- transform(lag.init, lower=mean-sem, upper=mean+sem)

# Yield
yield.init <- ddply(g.parms, c("Treatment"), summarise,
                   mean=mean(A), sem=sd(A)/sqrt(length(A)))
yield <- transform(yield.init, lower=mean-sem, upper=mean+sem)

# compare coefficients of variation of growth parameters as function of Rpf
g.parm.CV <- g.parms %>% group_by(Treatment) %>%
   summarise_at(c("L", "umax", "A"), cv)

# Relatively small difference in CV of yield for +Rpf (0.093) vs. -Rpf (0.131)
# CV for umax much higher in -Rpf (0.74) compared to +Rpf (0.0224)
# Similarly, CV for lag is higher in -Rpf (0.114) compared to +Rpf (0.0231)
```

# Statistically test for Rpf effect on growth curve parameters

```{r}
umax.test <- t.test(umax~Treatment, data=g.parms)
# t = 1.4039, df = 3.0013, p-value = 0.2549

lag.test <- t.test(L~Treatment, data=g.parms)
# t = 6.5252, df = 3.0966, p-value = 0.006631
# lag is 1.6 times or 65 % longer without Rpf
# Takes 179 h (or 7.5 days) longer to enter exponential phase

yield.test <- t.test(A~Treatment, data=g.parms)
# t = 1.6802, df = 5.0111, p-value = 0.1536
```

