---
title: "Rpf: Mutants"
author: "Jay T. Lennon and Venus Kuo"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

Termination of dormancy in soil bacteria with resuscitation promoting factor (Rpf)

## Clear workspace and set directory

```{r setup}
rm(list=ls())
getwd()
setwd("~/GitHub/Rpf")
```

## Load packages and functions

```{r}
require("png")
require("dplyr")
require("grid")
require("gdata")

sem <- function(x) sqrt(var(x)/length(x))
LL.95 <- function(x) t.test(x)$conf.int[1]
UL.95 <- function(x) t.test(x)$conf.int[2]
```

## Load growth data

Monitoring dormant KBS0714 populations with different mutations exposed to Rpf.
Mutations included substitution of conserved glutamic acid (E) catalytic site
with other amino acids, namely alanine (A), lysine (K), and gluatimne (Q). 
In addition, we truncated a repeating alanine residue in lectin-like regions, 
which may be involved in Rpf attachment to cell membrane. 

```{r}
mut <- read.delim("data/mutant/mutant.growth.txt", sep="\t", header = TRUE)
```

# Subset data for analyses and plotting
```{r}
E54A.data <- filter(mut, Experiment == "E54A") %>%
  dplyr::select(Treatment, OD600)
x1 <- c(6.5, 6.5, 6.5, 6.5, 4, 4, 4, 4, 1.5, 1.5, 1.5, 1.5)
E54A.data <- data.frame(E54A.data, x1)
E54A.data <- E54A.data[order(x1),]

E54A.summary <- filter(mut, Experiment == "E54A") %>%
  dplyr::select(OD600, Treatment) %>%
  group_by(Treatment) %>%
  summarize(mean = mean(OD600), SE = sem(OD600),
  LL.95 = LL.95(OD600), UL.95 = UL.95(OD600))
  x2 <- c(1.5, 6.5, 4)
  E54A.summary <- data.frame(E54A.summary,x2)
  E54A.summary <- E54A.summary[order(x2),]
  
E54A.anova <- aov(OD600 ~ Treatment, data = E54A.data)
E54A.tukey <- TukeyHSD(E54A.anova)

#             Df Sum Sq Mean Sq F value   Pr(>F)    
# Treatment    2  8.516   4.258   58.26 7.08e-06 ***
# Residuals    9  0.658   0.073                     
#
#                 diff        lwr        upr     p adj
# mut-control -0.65100 -1.1847231 -0.1172769 0.0192846
# wt-control   1.37025  0.8365269  1.9039731 0.0001392
# wt-mut       2.02125  1.4875269  2.5549731 0.0000060
 
# Compared to control, +Rpf WT increased: 95% or 1.95 fold (P = 0.0001)
# Compared to control, mutation decreased: 45% or 0.55 fold (P = 0.019)
# Compared to +Rpf WT, mutation decreased: 72% or 0.28 fold (P <0.0001)


E54K.data <- filter(mut, Experiment == "E54K") %>%
  dplyr::select(Treatment, OD600)
E54K.data <- data.frame(E54K.data, x1)
E54K.data <- E54K.data[order(x1),]

E54K.summary <- filter(mut, Experiment == "E54K") %>%
  dplyr::select(Treatment, OD600) %>%
  group_by(Treatment) %>%
  summarize(mean = mean(OD600), SE = sem(OD600),
  LL.95 = LL.95(OD600), UL.95 = UL.95(OD600))
  E54K.summary <- data.frame(E54K.summary,x2)
  E54K.summary <- E54K.summary[order(x2),]

E54K.anova <- aov(OD600 ~ Treatment, data = E54K.data)
E54K.tukey <- TukeyHSD(E54K.anova)

#             Df Sum Sq Mean Sq F value   Pr(>F)    
# Treatment    2  5.138   2.569   122.1 3.01e-07 ***
# Residuals    9  0.189   0.021   
#                diff       lwr       upr     p adj
# mut-control 0.47075 0.1843576 0.7571424 0.0033630
# wt-control  1.56225 1.2758576 1.8486424 0.0000003
# wt-mut      1.09150 0.8051076 1.3778924 0.0000057

# Compared to control, +Rpf WT increased: 128% or 2.28 fold (P < 0.0001)
# Compared to control, mutation increased: 39% or 1.39 fold (P = 0.003)
# Compared to +Rpf WT, mutation decreased: 39% or 0.61 fold (P < 0.0001)


E54Q.data <- filter(mut, Experiment == "E54Q") %>%
  dplyr::select(Treatment, OD600)
E54Q.data <- data.frame(E54Q.data, x1)
E54Q.data <- E54Q.data[order(x1),]

E54Q.summary <- filter(mut, Experiment == "E54Q") %>%
  dplyr::select(Treatment, OD600) %>%
  group_by(Treatment) %>%
  summarize(mean = mean(OD600), SE = sem(OD600),
  LL.95 = LL.95(OD600), UL.95 = UL.95(OD600))
  E54Q.summary <- data.frame(E54Q.summary,x2)
  E54Q.summary <- E54Q.summary[order(x2),]

E54Q.anova <- aov(OD600 ~ Treatment, data = E54Q.data)
E54Q.tukey <- TukeyHSD(E54Q.anova)

#            Df Sum Sq Mean Sq F value   Pr(>F)    
# Treatment    2  6.419   3.210   172.6 6.65e-08 ***
# Residuals    9  0.167   0.019 
#
#               diff       lwr       upr     p adj
# mut-control 1.17575 0.9065102 1.4449898 0.0000018
# wt-control  1.75850 1.4892602 2.0277398 0.0000001
# wt-mut      0.58275 0.3135102 0.8519898 0.0005038

# Compared to control, +Rpf WT increased: 299.5% or 4.0 fold (P < 0.0001)
# Compared to control, mutation increased: 200% or 3.0 fold (P < 0.0001)
# Compared to +Rpf WT, mutation decreased: 24.8% or 0.75 fold (P = 0.0005)

trnc.data <- filter(mut, Experiment == "trnc") %>%
  dplyr::select(Treatment, OD600)
x3 <- c(6.5, 6.5, 6.5, 6.5, 4, 4, 4, 4, 1.5, 1.5, 1.5)
trnc.data <- data.frame(trnc.data, x3)
trnc.data <- trnc.data[order(x3),]

trnc.summary <- filter(mut, Experiment == "trnc") %>%
  dplyr::select(Treatment, OD600) %>%
  group_by(Treatment) %>%
  summarize(mean = mean(OD600), SE = sem(OD600),
  LL.95 = LL.95(OD600), UL.95 = UL.95(OD600))
  trnc.summary <- data.frame(trnc.summary,x2)
  trnc.summary <- trnc.summary[order(x2),]

trnc.anova <- aov(OD600 ~ Treatment, data = trnc.data)
trnc.tukey <- TukeyHSD(trnc.anova)

#             Df Sum Sq Mean Sq F value   Pr(>F)    
# Treatment    2  3.917  1.9583   23.24 0.000465 ***
# Residuals    8  0.674  0.0843 
# 
#                  diff        lwr      upr     p adj
# mut-control 0.3965833 -0.2369275 1.030094 0.2335105
# wt-control  1.4235833  0.7900725 2.057094 0.0005271
# wt-mut      1.0270000  0.4404829 1.613517 0.0026580

# Compared to control, +Rpf WT increased: 273% or 3.7 fold (P = 0.0005)
# Compared to control, mutation increased: 76% or 1.8 fold (P = 0.234 --> NS)
# Compared to +Rpf WT, mutation decreased: 53% or 0.47 fold (P = 0.0003)
```

# Make figure

```{r}
png(filename="~/GitHub/Rpf/figures/Fig3.png",
    width = 1200, height = 1200, res = 96*2) 

plot.new()
par(oma=c(2,4,0,0), mar=c(2,4,4,1), mfrow=c(2,2))


# Panel A: E54A

# Add points
plot(jitter(E54A.data$x1, amount = 0.1), E54A.data$OD600, 
     ylim = c(-0.5, 5), xlim = c(0, 8), pch = 21, col = "lightgrey", 
     bg = "lightgrey", lwd = 2, cex = 1.7, yaxt = "n", xaxt = "n", cex.lab = 2, 
     cex.axis= 1.5, las = 1, ylab = "", xlab = "")
box(lwd=2) 

# Add means
points(E54A.summary$x2, E54A.summary$mean, pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)

# Add error bars
arrows(x0 = E54A.summary$x2, y0 = E54A.summary$mean, y1 = E54A.summary$LL.95, 
       angle = 90, length = 0.1, lwd = 2)
arrows(x0 = E54A.summary$x2, y0 = E54A.summary$mean, y1 = E54A.summary$UL.95, 
       angle = 90, length = 0.1, lwd = 2)

# Add axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("0.0", "2.0", "4.0"), at = c(0, 2, 4))
axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at=c(0, 2, 4), labels = F)
axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1.5, 4, 6.5), labels = F)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1.5, 4, 6.5), labels = F)

# Add x-axis labels
x.lab.cont <- list(bquote("-Rpf"))
mtext(do.call(expression, x.lab.cont),side = 1, 
          line = 2.4, at = 1.45, cex = 1, col = "black")

x.lab.wild <- list(bquote("+Rpf"))
mtext(do.call(expression, x.lab.wild),side = 1, 
          line = 2.4, at = 5.25, cex = 1, col = "black")

x.lab.wild.2 <- list(bquote("WT"))
mtext(do.call(expression, x.lab.wild.2), side = 1, 
          line = 0.75, at = 1.4, cex = 0.75, col = "gray40")

x.lab.wild.3 <- list(bquote("WT"))
mtext(do.call(expression, x.lab.wild.2), side = 1, 
          line = 0.75, at = 4.0, cex = 0.75, col = "gray40")

x.lab.mut.2 <- list(bquote(Delta[E %->% A]))
mtext(do.call(expression, x.lab.mut.2),side = 1, 
          line = 0.75, at = 6.5, cex = 0.75, col = "gray40")

# Add multiple comparisons
segments(1.5, 4.4, 6.5, 4.4, col = "black", lwd = 1.25, xpd = NA)
segments(1.5, 4.4, 1.5, 4.15, col = "black", lwd = 1.25, xpd = NA)
segments(6.5, 4.4, 6.5, 4.15, col = "black", lwd = 1.25, xpd = NA)
mtext(expression(~italic("P")~"= 0.019"), line = -1.6, cex = 0.5, at = 4)

segments(1.5, 3.6, 3.75, 3.6, col = "black", lwd = 1.25, xpd = NA)
segments(1.5, 3.6, 1.5, 3.35, col = "black", lwd = 1.25, xpd = NA)
segments(3.75, 3.6, 3.75, 3.35, col = "black", lwd = 1.25, xpd = NA)
mtext(expression(~italic("P")~"= 0.0001"), line = -3.2, cex = 0.5, at = 2.65)

segments(4.25, 3.6, 6.5, 3.6, col = "black", lwd = 1.25, xpd = NA)
segments(4.25, 3.6, 4.25, 3.35, col = "black", lwd = 1.25, xpd = NA)
segments(6.5, 3.6, 6.5, 3.35, col = "black", lwd = 1.25, xpd = NA)
mtext(expression(~italic("P")~"< 0.0001"), line = -3.2, cex = 0.5, at = 5.35)

# Rpf lines on x-axis label
segments(0.7, -1.7, 2.25, -1.7, col = "black", lwd = 2, xpd = TRUE)
segments(3.5, -1.7, 7.5, -1.7, col = "black", lwd = 2, xpd = TRUE)

# Add panel label
text(7.65, 4.7, labels = "A", cex = 2)

# Panel B: E54K

# Add points
plot(jitter(E54K.data$x1, amount = 0.1), E54K.data$OD600, 
     ylim = c(-0.5, 5), xlim = c(0, 8), pch = 21, col = "lightgrey", 
     bg = "lightgrey", lwd = 2, cex = 1.7, yaxt = "n", xaxt = "n", cex.lab = 2, 
     cex.axis= 1.5, las = 1, ylab = "", xlab = "")
box(lwd=2) 

# Add means
points(E54K.summary$x2, E54K.summary$mean, pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)

# Add error bars
arrows(x0 = E54K.summary$x2, y0 = E54K.summary$mean, y1 = E54K.summary$LL.95, 
       angle = 90, length = 0.1, lwd = 2)
arrows(x0 = E54K.summary$x2, y0 = E54K.summary$mean, y1 = E54K.summary$UL.95, 
       angle = 90, length = 0.1, lwd = 2)

# Add axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("0.0", "2.0", "4.0"), at = c(0, 2, 4))
axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at=c(0, 2, 4), labels = F)
axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1.5, 4, 6.5), labels = F)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1.5, 4, 6.5), labels = F)

# Add x-axis labels
x.lab.cont <- list(bquote("-Rpf"))
mtext(do.call(expression, x.lab.cont),side = 1, 
          line = 2.4, at = 1.45, cex = 1, col = "black")

x.lab.wild <- list(bquote("+Rpf"))
mtext(do.call(expression, x.lab.wild),side = 1, 
          line = 2.4, at = 5.25, cex = 1, col = "black")

x.lab.wild.2 <- list(bquote("WT"))
mtext(do.call(expression, x.lab.wild.2), side = 1, 
          line = 0.75, at = 1.4, cex = 0.75, col = "gray40")

x.lab.wild.3 <- list(bquote("WT"))
mtext(do.call(expression, x.lab.wild.2), side = 1, 
          line = 0.75, at = 4.0, cex = 0.75, col = "gray40")

x.lab.mut.2 <- list(bquote(Delta[E %->% K]))
mtext(do.call(expression, x.lab.mut.2),side = 1, 
          line = 0.75, at = 6.5, cex = 0.75, col = "gray40")

# Add multiple comparisons
segments(1.5, 4.4, 6.5, 4.4, col = "black", lwd = 1.25, xpd = NA)
segments(1.5, 4.4, 1.5, 4.15, col = "black", lwd = 1.25, xpd = NA)
segments(6.5, 4.4, 6.5, 4.15, col = "black", lwd = 1.25, xpd = NA)
mtext(expression(~italic("P")~"= 0.003"), line = -1.6, cex = 0.5, at = 4)

segments(1.5, 3.6, 3.75, 3.6, col = "black", lwd = 1.25, xpd = NA)
segments(1.5, 3.6, 1.5, 3.35, col = "black", lwd = 1.25, xpd = NA)
segments(3.75, 3.6, 3.75, 3.35, col = "black", lwd = 1.25, xpd = NA)
mtext(expression(~italic("P")~"< 0.0001"), line = -3.2, cex = 0.5, at = 2.75)

segments(4.25, 3.6, 6.5, 3.6, col = "black", lwd = 1.25, xpd = NA)
segments(4.25, 3.6, 4.25, 3.35, col = "black", lwd = 1.25, xpd = NA)
segments(6.5, 3.6, 6.5, 3.35, col = "black", lwd = 1.25, xpd = NA)
mtext(expression(~italic("P")~"< 0.0001"), line = -3.2, cex = 0.5, at = 5.25)

# Rpf lines on x-axis label
segments(0.7, -1.7, 2.25, -1.7, col = "black", lwd = 2, xpd = TRUE)
segments(3.5, -1.7, 7.5, -1.7, col = "black", lwd = 2, xpd = TRUE)

# Add panel label
text(7.65, 4.7, labels = "B", cex = 2)


# Panel C: E54Q

# Add points
plot(jitter(E54Q.data$x1, amount = 0.1), E54Q.data$OD600, 
     ylim = c(-0.5, 5), xlim = c(0.5, 8), pch = 21, col = "lightgrey", 
     bg = "lightgrey", lwd = 2, cex = 1.7, yaxt = "n", xaxt = "n", cex.lab = 2, 
     cex.axis= 1.5, las = 1, ylab = "", xlab = "")
box(lwd=2) 

# Add means
points(E54Q.summary$x2, E54Q.summary$mean, pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)

# Add error bars
arrows(x0 = E54Q.summary$x2, y0 = E54Q.summary$mean, y1 = E54Q.summary$LL.95, 
       angle = 90, length = 0.1, lwd = 2)
arrows(x0 = E54Q.summary$x2, y0 = E54Q.summary$mean, y1 = E54Q.summary$UL.95, 
       angle = 90, length = 0.1, lwd = 2)

# Add axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("0.0", "2.0", "4.0"), at = c(0, 2, 4))
axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at=c(0, 1, 2, 3), labels = F)
axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1.5, 4, 6.5), labels = F)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1.5, 4, 6.5), labels = F)

# Add x-axis labels

# Add x-axis labels
x.lab.cont <- list(bquote("-Rpf"))
mtext(do.call(expression, x.lab.cont),side = 1, 
          line = 2.4, at = 1.45, cex = 1, col = "black")

x.lab.wild <- list(bquote("+Rpf"))
mtext(do.call(expression, x.lab.wild),side = 1, 
          line = 2.4, at = 5.25, cex = 1, col = "black")

x.lab.wild.2 <- list(bquote("WT"))
mtext(do.call(expression, x.lab.wild.2), side = 1, 
          line = 0.75, at = 1.4, cex = 0.75, col = "gray40")

x.lab.wild.3 <- list(bquote("WT"))
mtext(do.call(expression, x.lab.wild.2), side = 1, 
          line = 0.75, at = 4.0, cex = 0.75, col = "gray40")

x.lab.mut.2 <- list(bquote(Delta[E %->% Q]))
mtext(do.call(expression, x.lab.mut.2),side = 1, 
          line = 0.75, at = 6.5, cex = 0.75, col = "gray40")


# Add multiple comparisons

segments(1.5, 4.4, 6.5, 4.4, col = "black", lwd = 1.25, xpd = NA)
segments(1.5, 4.4, 1.5, 4.15, col = "black", lwd = 1.25, xpd = NA)
segments(6.5, 4.4, 6.5, 4.15, col = "black", lwd = 1.25, xpd = NA)
mtext(expression(~italic("P")~"< 0.0001"), line = -1.6, cex = 0.5, at = 4)

segments(1.5, 3.6, 3.75, 3.6, col = "black", lwd = 1.25, xpd = NA)
segments(1.5, 3.6, 1.5, 3.35, col = "black", lwd = 1.25, xpd = NA)
segments(3.75, 3.6, 3.75, 3.35, col = "black", lwd = 1.25, xpd = NA)
mtext(expression(~italic("P")~"< 0.0001"), line = -3.2, cex = 0.5, at = 2.75)

segments(4.25, 3.6, 6.5, 3.6, col = "black", lwd = 1.25, xpd = NA)
segments(4.25, 3.6, 4.25, 3.35, col = "black", lwd = 1.25, xpd = NA)
segments(6.5, 3.6, 6.5, 3.35, col = "black", lwd = 1.25, xpd = NA)
mtext(expression(~italic("P")~"= 0.0005"), line = -3.2, cex = 0.5, at = 5.25)

# Rpf lines on x-axis label
segments(0.7, -1.7, 2.25, -1.7, col = "black", lwd = 2, xpd = TRUE)
segments(3.5, -1.7, 7.5, -1.7, col = "black", lwd = 2, xpd = TRUE)

# Add panel label
text(7.65, 4.7, labels = "C", cex = 2)


# Panel D: truncated

# Add points
plot(jitter(trnc.data$x3, amount = 0.1), trnc.data$OD600, 
     ylim = c(-0.5, 5), xlim = c(0.5, 8), pch = 21, col = "lightgrey", 
     bg = "lightgrey", lwd = 2, cex = 1.7, yaxt = "n", xaxt = "n", cex.lab = 2, 
     cex.axis= 1.5, las = 1, ylab = "", xlab = "")
box(lwd=2) 

# Add means
points(trnc.summary$x2, trnc.summary$mean, pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)

# Add error bars
arrows(x0 = trnc.summary$x2, y0 = trnc.summary$mean, y1 = trnc.summary$LL.95, 
       angle = 90, length = 0.1, lwd = 2)
arrows(x0 = trnc.summary$x2, y0 = trnc.summary$mean, y1 = trnc.summary$UL.95, 
       angle = 90, length = 0.1, lwd = 2)

# Add axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("0.0", "2.0", "4.0"), at = c(0, 2, 4))
axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at=c(0, 2, 4), labels = F)
axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1.5, 4, 6.5), labels = F)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1.5, 4, 6.5), labels = F)

# Add x-axis labels
x.lab.cont <- list(bquote("-Rpf"))
mtext(do.call(expression, x.lab.cont),side = 1, 
          line = 2.4, at = 1.45, cex = 1, col = "black")

x.lab.wild <- list(bquote("+Rpf"))
mtext(do.call(expression, x.lab.wild),side = 1, 
          line = 2.4, at = 5.25, cex = 1, col = "black")

x.lab.wild.2 <- list(bquote("WT"))
mtext(do.call(expression, x.lab.wild.2), side = 1, 
          line = 0.75, at = 1.4, cex = 0.75, col = "gray40")

x.lab.wild.3 <- list(bquote("WT"))
mtext(do.call(expression, x.lab.wild.2), side = 1, 
          line = 0.75, at = 4.0, cex = 0.75, col = "gray40")

x.lab.mut.2 <- list(bquote("Truncated"))
mtext(do.call(expression, x.lab.mut.2),side = 1, 
          line = 0.75, at = 6.5, cex = 0.75, col = "gray40")


# Add multiple comparisons

segments(1.5, 4.4, 6.5, 4.4, col = "black", lwd = 1.25, xpd = NA)
segments(1.5, 4.4, 1.5, 4.15, col = "black", lwd = 1.25, xpd = NA)
segments(6.5, 4.4, 6.5, 4.15, col = "black", lwd = 1.25, xpd = NA)
mtext(expression(~italic("P")~"= 0.234"), line = -1.6, cex = 0.5, at = 4)

segments(1.5, 3.6, 3.75, 3.6, col = "black", lwd = 1.25, xpd = NA)
segments(1.5, 3.6, 1.5, 3.35, col = "black", lwd = 1.25, xpd = NA)
segments(3.75, 3.6, 3.75, 3.35, col = "black", lwd = 1.25, xpd = NA)
mtext(expression(~italic("P")~"= 0.0005"), line = -3.2, cex = 0.5, at = 2.75)

segments(4.25, 3.6, 6.5, 3.6, col = "black", lwd = 1.25, xpd = NA)
segments(4.25, 3.6, 4.25, 3.35, col = "black", lwd = 1.25, xpd = NA)
segments(6.5, 3.6, 6.5, 3.35, col = "black", lwd = 1.25, xpd = NA)
mtext(expression(~italic("P")~"= 0.0003"), line = -3.2, cex = 0.5, at = 5.25)

# Add panel label
text(7.65, 4.7, labels = "D", cex = 2)

# Rpf lines on x-axis label
segments(0.7, -1.7, 2.25, -1.7, col = "black", lwd = 2, xpd = TRUE)
segments(3.5, -1.7, 7.7, -1.7, col = "black", lwd = 2, xpd = TRUE)

# Add x-axis labels
#mtext('-Rpf', side = 1, line = 4, at = 1, cex = 1.5)
#mtext('+Rpf', side = 1, line = 4, at = 3.5, cex = 1.5)

# Add outside y axis label
mtext(expression('Biomass (OD600)'), side = 2,
      outer = TRUE, cex = 2, line = 0.5, adj = 0.5)

# Close Plot Device
dev.off()
graphics.off()

img <- readPNG("./figures/Fig3.png")
grid.raster(img)
```