---
title: "Rpf: Antibody Controls"
author: "Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

Termination of dormancy in soil bacteria with resuscitation promoting factor (Rpf)

## Clear workspace and set directory

```{r setup}
rm(list=ls())
getwd()
setwd("~/GitHub/Rpf/")
```

## Load packages and functions

```{r}
require("png")
require("grid")
sem <- function(x) sqrt(var(x)/length(x))
```

## Confirmation of enzymatic mechanisms of resuscitation with antibody control

Measured abundance (OD600) of dormant KBS0714 under different conditions after
incubating at 25C and 200 rpm after 120 hrs:

1) No recombinant Rpf added
2) Plus recombiant Rpf (1.7 uM), but no antigen and no antibody
3) Plus recombinant Rpf (1.7 uM) and antibody, but no antigen 
4) Plus recombinant Rpf (1.7 uM), antibody, and antigen

For treatment 3, we added antibody taken at time zero before antigens developed
For treatment 4, we added antibody taken 28 days following injection 

## Load data and make a table

```{r}
antibody <- read.table("data/antibody/antibody.txt", header = TRUE, sep="\t")

rpf.mean <- aggregate(antibody$OD600 ~ Treatment, mean, data = antibody)
rpf.sem <- aggregate(antibody$OD600 ~ Treatment, sem, data = antibody)
LL.95 <- aggregate(antibody$OD600 ~ Treatment, antibody,
          FUN = function(x) t.test(x)$conf.int[1])
UL.95 <- aggregate(antibody$OD600 ~ Treatment, antibody,
          FUN = function(x) t.test(x)$conf.int[2])
table <- data.frame(rpf.mean[1], rpf.mean[2], rpf.sem[2],
          LL.95[2], UL.95[2])
colnames(table) <- c("Treatment", "mean", "sem", "LCI", "UCI")

# Rpf led to a 3.96-fold (or 296%) increase in growth
# Post-bleed compared to pre-bleed to 8.11-fold (or 88%) reduction
```

## Run an analysis of varaince (ANOVA)

```{r}
antibody.anova <- aov(OD600 ~ Treatment, data = antibody)
summary(antibody.anova)
TukeyHSD(antibody.anova)
  # no difference between -Rpf and +Rpf,+anti.28 --> P = 0.1166317
  # no difference between +Rpf and +Rpf,+anti.0 --> P = 0.9877646
  # for all other comparisons, P < 0.0001
```

## Make figure

```{r}
# Subset data
no.rpf <- antibody[ which(antibody$Treatment == "Rpf-"),]
plus.rpf <- antibody[ which(antibody$Treatment == "Rpf+"),]
plus.rpf.anti.0 <- antibody[ which(antibody$Treatment == "Rpf+, Anti0+"),]
plus.rpf.anti.28 <- antibody[ which(antibody$Treatment == "Rpf+, Anti28+"),]

# Create plot
png(filename="~/GitHub/Rpf/figures/FigS4.png",
    width = 1200, height = 1200, res = 96*2) 

plot.new()
par(mar = c(7, 7, 5, 7))

# Add individuals points
antibody.point <- plot(jitter(rep(1, length(no.rpf$OD600)), amount = 0.1), no.rpf$OD600, 
      ylim = c(-0.2, 3.7), xlim = c(0.5, 5), pch = 21, col = "lightgrey", bg = "lightgrey", lwd = 2,
      cex = 1.7, yaxt = "n", xaxt = "n", cex.lab = 2, cex.axis = 1.5,
      las = 1, ylab = "", xlab = "")
      box(lwd = 2)
      
points(jitter(rep(2.5, length(plus.rpf$OD600)), amount = 0.1), plus.rpf$OD600, pch = 21, 
       bg = "lightgrey", col = "lightgrey", lwd = 2, cex = 1.7)
points(jitter(rep(3.5, length(plus.rpf.anti.0$OD600)), amount = 0.1), plus.rpf.anti.0$OD600, pch = 21, 
       bg = "lightgrey", col = "lightgrey", lwd = 2, cex = 1.7)
points(jitter(rep(4.5, length(plus.rpf.anti.28$OD600)), amount = 0.1), plus.rpf.anti.28$OD600, pch = 21, 
       bg = "lightgrey", col = "lightgrey", lwd = 2, cex = 1.7)

# Add means
points(1, mean(no.rpf$OD600), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
points(2.5, mean(plus.rpf$OD600), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)  
points(3.5, mean(plus.rpf.anti.0$OD600), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
points(4.5, mean(plus.rpf.anti.28$OD600), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 

# Add error bars
arrows(x0 = c(1, 2.5, 3.5, 4.5), y0 = table$mean, y1 = table$LCI, angle = 90,
       length = 0.1, lwd = 2)
arrows(x0 = c(1, 2.5, 3.5, 4.5), y0 = table$mean, y1 = table$UCI, angle = 90,
       length=0.1, lwd = 2)

# Add axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("0.0", "1.0", "2.0", "3.0"), at = c(0, 1, 2, 3))

axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at=c(0, 1, 2, 3), labels = F)

axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1, 2.5, 3.5, 4.5), labels = F)

# Add y-axis label
mtext(expression('Biomass (OD600)'), side = 2,
      outer = FALSE, cex = 1.5, line = 3.6, adj = 0.5)

# Add x-axis labels
mtext('-Rpf', side = 1, line = 4, at = 1, cex = 1.5)
mtext('+Rpf', side = 1, line = 4, at = 3.5, cex = 1.5)

mtext('none', side = 1, line = 1.75, at = 2.5, cex = 0.8, col ='gray40')
#mtext('-', side = 1, line = 2.25, at = 2.5, cex = 2)
mtext('pre-bleed', side = 1, line = 1.75, at = 3.5, cex = 0.8, col ='gray40')
#mtext('+', side = 1, line = 0.75, at = 3.5, cex = 2)
#mtext('-', side = 1, line = 2.25, at = 3.5, cex = 2)
mtext('post-bleed', side = 1, line = 1.75, at = 4.5, cex = 0.8, col ='gray40')
#mtext('+', side = 1, line = 0.75, at = 4.5, cex = 2)
#mtext('+', side = 1, line = 2.25, at = 4.5, cex = 2)


mtext(expression(~italic('Antibody treatment')), side = 1, line = 0.7, 
      at = 3.5, cex = 0.95, col ='gray40')
#mtext('Antigen', side = 1, line = 2.25, at = 5.7, cex = 1.5)

segments(0.7, -1, 1.5, -1, col = "black", lwd = 2, xpd = TRUE)
segments(2.25, -1, 4.9, -1, col = "black", lwd = 2, xpd = TRUE)

text(1, 1.65, labels = "a", cex = 1.75, col = 'gray60')
text(2.5, 3.4, labels = "b", cex = 1.75, col = 'gray60')
text(3.5, 3.4, labels = "b", cex = 1.75, col = 'gray60')
text(4.5, 1.0, labels = "a", cex = 1.75, col = 'gray60')

# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("./figures/FigS4.png")

grid.raster(img)
```
