---
title: "Rpf"
author: "Jay T. Lennon and Venus Kuo"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

Generality of Rpf on KBS bacterial isolates

# Set up working environment

```{r setup}
rm(list=ls())
getwd()
#setwd("~/GitHub/Rpf")
knitr::opts_knit$set(root.dir = '~/../GitHub/Rpf') 

sem <- function(x) sqrt(var(x)/length(x))
```

# Load data set

```{r}
# A) Generate % change in growth curve parameter with Rpf treatment among KBS0714 strains # 
# Load gompertz output data 
gc.param.1 <- read.delim("data/generality2/20180417_generality2_p1_gomp_output.txt", sep = ",", header = TRUE)
gc.param.2 <- read.delim("data/generality2/20180419_generality2_p2_gomp_output.txt", sep = ",", header = TRUE)
gc.param.3 <- read.delim("data/generality2/20180430_generality2_p3_gomp_output.txt", sep = ",", header = TRUE)

# Load key for growth curve assay 
gc.key <- read.delim("data/generality2/generality2_key.txt", sep = "\t", header = TRUE)

# Label phylogenetic and treatment information in datasets 
key.1 <- subset(gc.key, Run == "1", select = c(Strain,Gram,Phylum,Genus,Treatment,Rep))
gc.param.1 <- cbind(gc.param.1,key.1)
key.2 <- subset(gc.key, Run == "2", select = c(Strain,Gram,Phylum,Genus,Treatment,Rep))
gc.param.2 <- cbind(gc.param.2,key.2)
key.3 <- subset(gc.key, Run == "3", select = c(Strain,Gram,Phylum,Genus,Treatment,Rep))
gc.param.3 <- cbind(gc.param.3,key.3)

# Merge datasets
gc <- rbind(gc.param.1,gc.param.2, gc.param.3)

# Remove unused parameters to simplify
gc <- gc[ ,c(-(1:2),-(6:10))]

# Some biomass yeild values are negative, force to be positive # 
gc$A <- abs(gc$A)

# Remove sample 1, 3, and 15 - cultures never took off
gc <- gc[-c(1,3,15), ]

```

# Plot bargraph to look at growth curve parameter results

```{r}
# A) Generate simple bargraph to look at the effects of Rpf on growth curve parameters
library(ggplot2)
library(plyr)
# Calculate mean and sem 
# Biomass yeild (A)
generality.A.means.sem <- ddply(gc, c("Genus","Treatment"), summarise,
                   mean=mean(A), sem=sd(A)/sqrt(length(A)))
generality.A.means.sem <- transform(generality.A.means.sem, lower=mean-sem, upper=mean+sem)

generality.A.means.sem$Genus <- factor(generality.A.means.sem$Genus,
                                       levels = c("Micrococcus", "Arthrobacter2", "Curtobacterium", 
                                                  "Mycobacterium", "Bacillus","Janthinobacterium", "Pedobacter", "Pseudomonas"))
# Maximum growth rate (umax)
generality.umax.means.sem <- ddply(gc, c("Genus", "Treatment"), summarise,
                   mean=mean(umax), sem=sd(umax)/sqrt(length(umax)))
generality.umax.means.sem <- transform(generality.umax.means.sem, lower=mean-sem, upper=mean+sem)

generality.umax.means.sem$Genus <- factor(generality.umax.means.sem$Genus,
                                       levels = c("Micrococcus", "Arthrobacter2", "Curtobacterium", 
                                                  "Mycobacterium", "Bacillus","Janthinobacterium", "Pedobacter", "Pseudomonas"))
# Lag time (L)
generality.L.means.sem <- ddply(gc, c("Genus","Treatment"), summarise,
                   mean=mean(L), sem=sd(L)/sqrt(length(L)))
generality.L.means.sem <- transform(generality.L.means.sem, lower=mean-sem, upper=mean+sem)
generality.L.means.sem$Genus <- factor(generality.L.means.sem$Genus,
                                       levels = c("Micrococcus", "Arthrobacter2", "Curtobacterium", 
                                                  "Mycobacterium", "Bacillus","Janthinobacterium", "Pedobacter", "Pseudomonas"))

# Bargraph figure 
A.bar <- ggplot(data=generality.A.means.sem, aes(x=Genus, y=mean, fill=Treatment)) + 
  geom_bar(colour="black", position = position_dodge(), stat = "identity", size = 1)

A.bar + geom_errorbar(aes(ymax=upper, ymin=lower), position=position_dodge(1), 
                          data=generality.A.means.sem, width = 0.35, size=1) +
  labs(x =  "Genus" ,y="Biomass Yield") +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size = 12),
        axis.text.x=element_text(colour="black", size=12, angle =45,hjust = 1),
        axis.title = element_text(size=18),
        axis.title.y = element_text(colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(colour="black",margin=margin(15,10,0,10)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2))

umax.bar <- ggplot(data=generality.umax.means.sem, aes(x=Genus, y=mean, fill=Treatment)) + 
  geom_bar(colour="black", position = position_dodge(), stat = "identity", size = 1)

umax.bar + geom_errorbar(aes(ymax=upper, ymin=lower), position=position_dodge(1), 
                          data=generality.umax.means.sem, width = 0.35, size=1) +
  labs(x =  "Genus" ,y="Maximum Growth Rate") +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size = 12),
        axis.text.x=element_text(colour="black", size=12, angle =45,hjust = 1),
        axis.title = element_text(size=18),
        axis.title.y = element_text(colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(colour="black",margin=margin(15,10,0,10)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2))

L.bar <- ggplot(data=generality.L.means.sem, aes(x=Genus, y=mean, fill=Treatment)) + 
  geom_bar(colour="black", position = position_dodge(), stat = "identity", size = 1)

L.bar + geom_errorbar(aes(ymax=upper, ymin=lower), position=position_dodge(1), 
                          data=generality.L.means.sem, width = 0.35, size=1) +
  labs(x =  "Genus" ,y="Lag Time") +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size = 12),
        axis.text.x=element_text(colour="black", size=12, angle =45,hjust = 1),
        axis.title = element_text(size=18),
        axis.title.y = element_text(colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(colour="black",margin=margin(15,10,0,10)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2))
# Rpf+ treatment appears to only have a major effect on decrease the L and increasing the A of Micrococcus. 
# The other distantly related KBS strains did not appear to respond in significant ways to Rpf+ treatment. 
```

# Statistical analysis of KBS growth curve response to Rpf treatment by strain and gram status 

```{r}
# Does rpf effect on growth parameters vary among species (ANOVA)
# two-factor ANOVA
generality.umax.anova <- aov(umax ~ Treatment*Genus, data = gc)
summary(generality.umax.anova)

generality.L.anova <- aov(L ~ Treatment*Genus, data = gc)
summary(generality.L.anova)

generality.A.anova <- aov(A ~ Treatment*Genus, data = gc)
summary(generality.A.anova)

# For Micrococcus, Baccilus, and Janthino strains, Rpf treatment had a significant effect on A and L but this pattern is driven by Micrococcus response to Rpf+

# Statistical tests: Are effects influenced by whether strain is Gram + vs. Gram -?
# two-factor ANOVA
generality.g.umax.anova <- aov(umax ~ Treatment*Gram, data = gc)
summary(generality.g.umax.anova)

generality.g.L.anova <- aov(L ~ Treatment*Gram, data = gc)
summary(generality.g.L.anova)

generality.g.A.anova <- aov(A ~ Treatment*Gram, data = gc)
summary(generality.g.A.anova)
# Currently not enought samples
```

# Calculate proportional change in growth curve parameter

```{r}
# Calculate % Biomass yeild (A) change
a <- as.data.frame(gc$A[which(gc$Treatment == "Rpf+")])
b <- as.data.frame(gc$A[which(gc$Treatment == "Rpf-")])
df <- c()
df <- as.data.frame(((a-b)/b)*100)
# Calculate % Maximum growth rate (umax) change
c <- as.data.frame(gc$umax[which(gc$Treatment == "Rpf+")])
d <- as.data.frame(gc$umax[which(gc$Treatment == "Rpf-")])
df.u <- as.data.frame(((c-d)/d)*100)
df <- cbind(df, df.u)
# Calculate % Lag time (L) change
e <- gc$L[which(gc$Treatment == "Rpf+")]
f <- gc$L[which(gc$Treatment == "Rpf-")]
df.l <- as.data.frame(((e-f)/f)*100)
df <- cbind(df, df.l)
colnames(df) <- c("A", "umax", "L")

# Add phylogenetic information to dataset
df.info <- as.data.frame(gc[c(1:10, 21:32, 45:52), 4:7])
df <- cbind(df, df.info) 

# Non-normalizing 
#df.norm <- df

# Normalizing data
df.norm <- as.data.frame(scale(df[,1:3]))
df.norm <- cbind(df.norm, df[, 4:7])

# Generate L table of means, sems, and CIs 
df.means <- aggregate(L ~ Genus, mean, data= df.norm)
df.sems <- aggregate(L ~ Genus, sem, data = df.norm)
df.LL.95 <- aggregate(df.norm$L ~ Genus, df.norm,
          FUN = function(x) t.test(x)$conf.int[1])
df.UL.95 <- aggregate(df.norm$L ~ Genus, df.norm,
          FUN = function(x) t.test(x)$conf.int[2])
df.table <- data.frame(df.means[1], df.means[2], df.sems[2],
                       df.LL.95[2], df.UL.95[2])
colnames(df.table) <- c( "Genus",  "mean", "sem", "LCI", "UCI")

# Order by PD
label <- c(2,5,3,6,1,4,8,7)
df.table$PD <- label
df.table <- df.table[order(df.table$PD), ]


# Generate A table of means, sems, and CIs
df.A.means <- aggregate(A ~ Genus, mean, data= df.norm)
df.A.sems <- aggregate(A ~ Genus, sem, data = df.norm)
df.A.LL.95 <- aggregate(df.norm$A ~ Genus, df.norm,
          FUN = function(x) t.test(x)$conf.int[1])
df.A.UL.95 <- aggregate(df.norm$A ~ Genus, df.norm,
          FUN = function(x) t.test(x)$conf.int[2])
df.A.table <- data.frame(df.A.means[1], df.A.means[2], df.A.sems[2],
                       df.A.LL.95[2], df.A.UL.95[2])
colnames(df.A.table) <- c( "Genus",  "mean", "sem", "LCI", "UCI")

# Order by PD
df.A.table$PD <- label
df.A.table <- df.A.table[order(df.A.table$PD), ]

# Generate umax table of means, sems, and CIs
df.umax.means <- aggregate(umax ~ Genus, mean, data= df.norm)
df.umax.sems <- aggregate(umax ~ Genus, sem, data = df.norm)
df.umax.LL.95 <- aggregate(df.norm$umax ~ Genus, df.norm,
          FUN = function(x) t.test(x)$conf.int[1])
df.umax.UL.95 <- aggregate(df.norm$umax ~ Genus, df.norm,
          FUN = function(x) t.test(x)$conf.int[2])
df.umax.table <- data.frame(df.umax.means[1], df.umax.means[2], df.umax.sems[2],
                       df.umax.LL.95[2], df.umax.UL.95[2])
colnames(df.umax.table) <- c( "Genus",  "mean", "sem", "LCI", "UCI")

# Order by PD
df.umax.table$PD <- label
df.umax.table <- df.umax.table[order(df.umax.table$PD), ]

# Subset data for figure 
df.norm.Ped <- df.norm[ which(df.norm$Genus == "Pedobacter"), ]
#df.norm.Art1 <- df.norm[ which(df.norm$Genus == "Arthrobacter1"), ]
df.norm.Art2 <- df.norm[ which(df.norm$Genus == "Arthrobacter2"), ]
df.norm.Myc <- df.norm[ which(df.norm$Genus == "Mycobacterium"), ]
#df.norm.Azo <- df.norm[ which(df.norm$Genus == "Azospirillium"), ]
#df.norm.Rho <- df.norm[ which(df.norm$Genus == "Rhodococcus"), ]
df.norm.Pse <- df.norm[ which(df.norm$Genus == "Pseudomonas"), ]
df.norm.Jan <- df.norm[ which(df.norm$Genus == "Janthinobacterium"), ]
#df.norm.Var <- df.norm[ which(df.norm$Genus == "Variovorax"), ]
df.norm.Mic <- df.norm[ which(df.norm$Genus == "Micrococcus"), ]
df.norm.Cur <- df.norm[ which(df.norm$Genus == "Curtobacterium"), ]
df.norm.Bac <- df.norm[ which(df.norm$Genus == "Bacillus"), ]
```

# Visualizing % Change without normalization 

```{r}
df.table$Genus <- factor(df.table$Genus,levels = c("Micrococcus", "Arthrobacter2", "Curtobacterium", 
                                                  "Mycobacterium", "Bacillus","Janthinobacterium"))
df.A.table$Genus <- factor(df.A.table$Genus,levels = c("Micrococcus", "Arthrobacter2", "Curtobacterium", 
                                                  "Mycobacterium", "Bacillus","Janthinobacterium"))
df.umax.table$Genus <- factor(df.umax.table$Genus,levels = c("Micrococcus", "Arthrobacter2", "Curtobacterium", 
                                                  "Mycobacterium", "Bacillus","Janthinobacterium"))

df.table <- transform(df.table, lower=mean-sem, upper=mean+sem)
df.A.table <- transform(df.A.table, lower=mean-sem, upper=mean+sem)
df.umax.table <- transform(df.umax.table, lower=mean-sem, upper=mean+sem)


ggplot(data=df.table, aes(x=Genus, y=mean)) + 
  geom_bar(colour="black", position = position_dodge(), stat = "identity", size = 1) + 
  labs(x =  "Genus" ,y="Lag Time") +
  geom_errorbar(aes(ymax=upper, ymin=lower), position=position_dodge(1), 
                          data=df.table, width = 0.35, size=1) +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size = 12),
        axis.text.x=element_text(colour="black", size=12, angle =45,hjust = 1),
        axis.title = element_text(size=18),
        axis.title.y = element_text(colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(colour="black",margin=margin(15,10,0,10)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2))

ggplot(data=df.A.table, aes(x=Genus, y=mean)) + 
  geom_bar(colour="black", position = position_dodge(), stat = "identity", size = 1) +
  labs(x = "Genus", y = "Biomass yield")+
  geom_errorbar(aes(ymax=upper, ymin=lower), position=position_dodge(1), 
                          data=df.A.table, width = 0.35, size=1) +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size = 12),
        axis.text.x=element_text(colour="black", size=12, angle =45,hjust = 1),
        axis.title = element_text(size=18),
        axis.title.y = element_text(colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(colour="black",margin=margin(15,10,0,10)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2))

ggplot(data=df.umax.table, aes(x=Genus, y=mean)) + 
  geom_bar(colour="black", position = position_dodge(), stat = "identity", size = 1) + 
  labs(x = "Genus", y = "Maximum Growth rate")+
  geom_errorbar(aes(ymax=upper, ymin=lower), position=position_dodge(1), 
                          data=df.umax.table, width = 0.35, size=1) +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size = 12),
        axis.text.x=element_text(colour="black", size=12, angle =45,hjust = 1),
        axis.title = element_text(size=18),
        axis.title.y = element_text(colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(colour="black",margin=margin(15,10,0,10)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2))
```


# Clean figure with 95% CI with normalization

```{r}
require("png")
require("grid")

# Generate Alt. 3x1 figure
png(filename="~/../GitHub/Rpf/figures/Fig7.png",  
    width = 1200, height = 1200, res = 96*2)
plot.new()
par(oma=c(3,5,0,0), mar=c(1,2,1,1), mfrow = c(3,1),  byrow = TRUE)

# Lag time
plot(jitter(rep(1, length(df.norm.Mic$L)), amount = 0.1), df.norm.Mic$L,
     ylim = c(-3, 3), xlim = c(0.5, 8.5), pch = 21, col = "grey", bg = "grey", lwd = 2,
           cex = 1.7, yaxt = "n", xaxt = "n", cex.lab = 2, cex.axis = 1.5,
      las = 1, ylab = "", xlab = "")
      box(lwd = 2)
#points(jitter(rep(2, length(df.norm.Art1$L)), amount = 0.1), df.norm.Art1$L, pch = 21, 
#       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(2, length(df.norm.Art2$L)), amount = 0.1), df.norm.Art2$L, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(3, length(df.norm.Cur$L)), amount = 0.1), df.norm.Cur$L, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
#points(jitter(rep(4, length(df.norm.Rho$L)), amount = 0.1), df.norm.Rho$L, pch = 21, 
#       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(4, length(df.norm.Myc$L)), amount = 0.1), df.norm.Myc$L, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(5, length(df.norm.Bac$L)), amount = 0.1), df.norm.Bac$L, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
#points(jitter(rep(7, length(df.norm.Azo$L)), amount = 0.1), df.norm.Azo$L, pch = 21, 
#       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(6, length(df.norm.Jan$L)), amount = 0.1), df.norm.Jan$L, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(7, length(df.norm.Pse$L)), amount = 0.1), df.norm.Pse$L, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
#points(jitter(rep(10, length(df.norm.Var$L)), amount = 0.1), df.norm.Var$L, pch = 21, 
#       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(8, length(df.norm.Ped$L)), amount = 0.1), df.norm.Ped$L, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)

points(1, mean(df.norm.Mic$L), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
#points(2, mean(df.norm.Art1$L), pch = 21, col = "black", 
#       bg = "NA", lwd = 2, cex = 2.5)  
points(2, mean(df.norm.Art2$L), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
points(3, mean(df.norm.Cur$L), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
#points(4, mean(df.norm.Rho$L), pch = 21, col = "black", 
#       bg = "NA", lwd = 2, cex = 2.5) 
points(4, mean(df.norm.Myc$L), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)  
points(5, mean(df.norm.Bac$L), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
#points(7, mean(df.norm.Azo$L), pch = 21, col = "black", 
#       bg = "NA", lwd = 2, cex = 2.5) 
points(6, mean(df.norm.Jan$L), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
points(7, mean(df.norm.Pse$L), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
#points(10, mean(df.norm.Var$L), pch = 21, col = "black", 
#       bg = "NA", lwd = 2, cex = 2.5) 
points(8, mean(df.norm.Ped$L), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 

# Y-axis label
mtext(expression('Lag Time (L)'), side = 2,
      outer = FALSE, cex = 1.25, line = 3, adj = 0.5)

# Major Axes
#axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
#     labels = F, at = c(1, 2, 3,4,5,6,7,8,9,10,11), las =3)

axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = F, at = c(1, 2, 3,4,5,6,7,8), las =3)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("-3.0", "-2.0", "-1.0", "0.0", "1.0", "2.0", "3.0"), at = c(-3, -2, -1, 0, 1, 2, 3))

axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(-3, -2, -1, 0, 1, 2, 3), labels = F)

#axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
#     at = c(1, 2, 3,4,5,6,7,8,9,10,11), labels = F)

axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1, 2, 3,4,5,6,7,8), labels = F)

# Add dotted line at zero
abline(h=0, col="black", lty =2, lwd=2)

#arrows(x0 = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11), y0 = df.table$mean, y1 = df.table$LCI, angle = 90,
#       length = 0.1, lwd = 2)
#arrows(x0 = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11), y0 = df.table$mean, y1 = df.table$UCI, angle = 90,
#       length=0.1, lwd = 2)

arrows(x0 = c(1, 2, 3,4,5,6,7,8), y0 = df.table$mean, y1 = df.table$LCI, angle = 90,
       length = 0.1, lwd = 2)
arrows(x0 = c(1, 2, 3,4,5,6,7,8), y0 = df.table$mean, y1 = df.table$UCI, angle = 90,
       length=0.1, lwd = 2)

box(lwd = 2)


# Biomass Yield
plot(jitter(rep(1, length(df.norm.Mic$A)), amount = 0.1), df.norm.Mic$A,
     ylim = c(-3, 3), xlim = c(0.5, 8.5), pch = 21, col = "grey", bg = "grey", lwd = 2,
           cex = 1.7, yaxt = "n", xaxt = "n", cex.lab = 2, cex.axis = 1.5,
      las = 1, ylab = "", xlab = "")
      box(lwd = 2)
#points(jitter(rep(2, length(df.norm.Art1$A)), amount = 0.1), df.norm.Art1$A, pch = 21, 
#       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(2, length(df.norm.Art2$A)), amount = 0.1), df.norm.Art2$A, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(3, length(df.norm.Cur$A)), amount = 0.1), df.norm.Cur$A, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
#points(jitter(rep(4, length(df.norm.Rho$A)), amount = 0.1), df.norm.Rho$A, pch = 21, 
#       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(4, length(df.norm.Myc$A)), amount = 0.1), df.norm.Myc$A, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(5, length(df.norm.Bac$A)), amount = 0.1), df.norm.Bac$A, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
#points(jitter(rep(7, length(df.norm.Azo$A)), amount = 0.1), df.norm.Azo$A, pch = 21, 
#       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(6, length(df.norm.Jan$A)), amount = 0.1), df.norm.Jan$A, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(7, length(df.norm.Pse$A)), amount = 0.1), df.norm.Pse$A, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
#points(jitter(rep(10, length(df.norm.Var$A)), amount = 0.1), df.norm.Var$A, pch = 21, 
#       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(8, length(df.norm.Ped$A)), amount = 0.1), df.norm.Ped$A, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)

points(1, mean(df.norm.Mic$A), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
#points(2, mean(df.norm.Art1$A), pch = 21, col = "black", 
#       bg = "NA", lwd = 2, cex = 2.5)  
points(2, mean(df.norm.Art2$A), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
points(3, mean(df.norm.Cur$A), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
#points(4, mean(df.norm.Rho$A), pch = 21, col = "black", 
#       bg = "NA", lwd = 2, cex = 2.5) 
points(4, mean(df.norm.Myc$A), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)  
points(5, mean(df.norm.Bac$A), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
#points(7, mean(df.norm.Azo$A), pch = 21, col = "black", 
#       bg = "NA", lwd = 2, cex = 2.5) 
points(6, mean(df.norm.Jan$A), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
points(7, mean(df.norm.Pse$A), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
#points(10, mean(df.norm.Var$A), pch = 21, col = "black", 
#       bg = "NA", lwd = 2, cex = 2.5) 
points(8, mean(df.norm.Ped$A), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 

# Outer Y-axis label
mtext(expression('Normalized % Change with Rpf Treatment'), side = 2,
      outer = TRUE, cex = 1.5, line = 3, adj = 0.5)
# Y-axis label
mtext(expression('Biomass Yield (A)'), side = 2,
      outer = FALSE, cex = 1.25, line = 3, adj = 0.5)

# Major Axes
#axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
#     labels =  FALSE, at = c(1, 2, 3,4,5,6,7,8,9,10,11), las =3)

axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = F, at = c(1, 2, 3,4,5,6,7,8), las =3)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("-3.0", "-2.0", "-1.0", "0.0", "1.0", "2.0", "3.0"), at = c(-3, -2, -1, 0, 1, 2, 3))

axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(-3, -2, -1, 0, 1, 2, 3), labels = F)

#axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
#     at = c(1, 2, 3,4,5,6,7,8,9,10,11), labels = F)

axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1, 2, 3, 4, 5, 6,7,8), labels = F)

# Add dotted line at zero
abline(h=0, col="black", lty =2, lwd=2)

arrows(x0 = c(1, 2, 3, 4, 5, 6,7,8), y0 = df.A.table$mean, y1 = df.A.table$LCI, angle = 90,
       length = 0.1, lwd = 2)
arrows(x0 = c(1, 2, 3, 4, 5, 6,7,8), y0 = df.A.table$mean, y1 = df.A.table$UCI, angle = 90,
       length=0.1, lwd = 2)

box(lwd = 2)


# Maximum growth rate (umax)
plot(jitter(rep(1, length(df.norm.Mic$umax)), amount = 0.1), df.norm.Mic$umax,
     ylim = c(-3, 3.5), xlim = c(0.5, 8.5), pch = 21, col = "grey", bg = "grey", lwd = 2,
           cex = 1.7, yaxt = "n", xaxt = "n", cex.lab = 2, cex.axis = 1.5,
      las = 1, ylab = "", xlab = "")
      box(lwd = 2)
#points(jitter(rep(2, length(df.norm.Art1$A)), amount = 0.1), df.norm.Art1$A, pch = 21, 
#       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(2, length(df.norm.Art2$umax)), amount = 0.1), df.norm.Art2$umax, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(3, length(df.norm.Cur$umax)), amount = 0.1), df.norm.Cur$umax, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
#points(jitter(rep(4, length(df.norm.Rho$umax)), amount = 0.1), df.norm.Rho$umax, pch = 21, 
#       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(4, length(df.norm.Myc$umax)), amount = 0.1), df.norm.Myc$umax, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(5, length(df.norm.Bac$umax)), amount = 0.1), df.norm.Bac$umax, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
#points(jitter(rep(7, length(df.norm.Azo$umax)), amount = 0.1), df.norm.Azo$umax, pch = 21, 
#       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(6, length(df.norm.Jan$umax)), amount = 0.1), df.norm.Jan$umax, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(7, length(df.norm.Pse$umax)), amount = 0.1), df.norm.Pse$umax, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
#points(jitter(rep(10, length(df.norm.Var$umax)), amount = 0.1), df.norm.Var$umax, pch = 21, 
#       bg = "grey", col = "grey", lwd = 2, cex = 1.7)
points(jitter(rep(8, length(df.norm.Ped$umax)), amount = 0.1), df.norm.Ped$umax, pch = 21, 
       bg = "grey", col = "grey", lwd = 2, cex = 1.7)

points(1, mean(df.norm.Mic$umax), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
#points(2, mean(df.norm.Art1$umax), pch = 21, col = "black", 
#       bg = "NA", lwd = 2, cex = 2.5)  
points(2, mean(df.norm.Art2$umax), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
points(3, mean(df.norm.Cur$umax), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
#points(4, mean(df.norm.Rho$umax), pch = 21, col = "black", 
#       bg = "NA", lwd = 2, cex = 2.5) 
points(4, mean(df.norm.Myc$umax), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)  
points(5, mean(df.norm.Bac$umax), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
#points(7, mean(df.norm.Azo$umax), pch = 21, col = "black", 
#       bg = "NA", lwd = 2, cex = 2.5) 
points(6, mean(df.norm.Jan$umax), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
points(7, mean(df.norm.Pse$umax), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
#points(10, mean(df.norm.Var$umax), pch = 21, col = "black", 
#       bg = "NA", lwd = 2, cex = 2.5) 
points(8, mean(df.norm.Ped$umax), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 

# Y-axis label
mtext(expression('Umax'), side = 2,
      outer = FALSE, cex = 1.25, line = 3, adj = 0.5)

# Major Axes
#axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
#     labels = c("Mic", "Art2", "Cur", "Rho", "Myc", "Bac", "Azo", "Jan", "Pse", "Var", "Ped"), 
#     at = c(1, 2, 3,4,5,6,7,8,9,10,11), las =3)

axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("Mic", "Art2", "Cur", "Myc", "Bac", "Jan", "Pse", "Ped"), 
     at = c(1, 2, 3,4,5,6,7,8), las =3)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("-3.0", "-2.0", "-1.0", "0.0", "1.0", "2.0", "3.0"), at = c(-3, -2, -1, 0, 1, 2, 3))

axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(-3, -2, -1, 0, 1, 2, 3), labels = F)

axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1, 2, 3,4,5,6,7,8), labels = F)

# Add dotted line at zero
abline(h=0, col="black", lty =2, lwd=2)

arrows(x0 = c(1, 2, 3 , 4, 5, 6,7,8), y0 = df.umax.table$mean, y1 = df.umax.table$LCI, angle = 90,
       length = 0.1, lwd = 2)
arrows(x0 = c(1, 2, 3, 4, 5, 6,7,8), y0 = df.umax.table$mean, y1 = df.umax.table$UCI, angle = 90,
       length=0.1, lwd = 2)

box(lwd = 2)

# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("~/../GitHub/Rpf/figures/Fig7.png")
grid.raster(img)
```

# Analysis of CFU counts from Generality Experiment 

```{r}
colony <- read.delim("data/generality2/Rpf_Generality_CFU_p1.txt", sep = ",", header = TRUE)

# Make Strains as factor
colony$Strain <- as.factor(colony$Strain)

# Calculate CFU
colony$cfu <- (colony$Colony*colony$Dilution)/colony$Incolumn_mL

# Plot results
cfu.means.sem <- ddply(colony, c("Strain","Treatment"), summarise,
                   mean=mean(cfu), sem=sd(cfu)/sqrt(length(cfu)))
cfu.means.sem <- transform(cfu.means.sem, lower=mean-sem, upper=mean+sem)


# Bargraph figure 
A.bar <- ggplot(data=cfu.means.sem, aes(x=Strain, y=mean, fill=Treatment)) + 
  geom_bar(colour="black", position = position_dodge(), stat = "identity", size = 1)

A.bar + geom_errorbar(aes(ymax=upper, ymin=lower), position=position_dodge(1), 
                          data=cfu.means.sem, width = 0.35, size=1) +
  labs(x =  "Strains" ,y="CFU") +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size = 12),
        axis.text.x=element_text(colour="black", size=12, angle =45,hjust = 1),
        axis.title = element_text(size=18),
        axis.title.y = element_text(colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(colour="black",margin=margin(15,10,0,10)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2))
```

# Statistical test of CFU 

```{r}
KBS812.cfu <- t.test(cfu ~ Treatment, data=colony[which(colony$Strain == "812"), ])
KBS812.cfu

KBS711.cfu <- t.test(cfu ~ Treatment, data=colony[which(colony$Strain == "711"), ])
KBS711.cfu

# No significant difference in CFU counts between treatments of strains
```

















