---
title: "Rpf: Growth Curve"
author: "Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

Termination of dormancy in soil bacteria with resuscitation promoting factor (Rpf)

## Clear workspace and set directory

```{r setup}
rm(list=ls())
getwd()
setwd("~/GitHub/Rpf")
```

## Load packages and functions

```{r}
require("png")
require("grid")
require("plyr")
require("dplyr")
sem <- function(x) sqrt(var(x)/length(x))
cv <- function(x) sd(x)/mean(x)
LL.95 <- function(x) t.test(x)$conf.int[1]
UL.95 <- function(x) t.test(x)$conf.int[2]
```

## Load growth data

This included monitoring of dormant KBS0714 populations in Klett flasks. 
Measured OD600 daily on flasks with (478 nM) vs. without (protein bufffer) Rpf.

```{r}
growth <- read.table("data/growth.curve/growth.curve.txt", sep="\t", header=TRUE)
```

## Make figure - time series

```{r}
png(filename="~/GitHub/Rpf/figures/Fig4.png",
    width = 1200, height = 900, res = 96*2) 
plot.new()
par(mar = c(7, 7, 5, 7))
# Add +Rpf tubes
plot(growth[,1], growth[,6], xlim = c(-25, 625), ylim = c(-0.5, 4), type = "l", 
     lty = 2, col = "black", lwd = 2, ylab = "", xlab = "", 
     cex.lab = 1.5, las = 1, yaxt = "n", xaxt = "n")
box(lwd = 2)
points(growth[,1], growth[,7], type = "l", lty = 2, col = "black", lwd = 2)
points(growth[,1], growth[,8], type = "l", lty = 2, col = "black", lwd = 2)
points(growth[,1], growth[,9], type = "l", lty = 2, col = "black", lwd = 2)
# Add -Rpf tubes
points(growth[,1], growth[,2], type = "l", lty = 3, col = "black", lwd = 2)
points(growth[,1], growth[,3], type = "l", lty = 3, col = "black", lwd = 2)
points(growth[,1], growth[,4], type = "l", lty = 3, col = "black", lwd = 2)
points(growth[,1], growth[,5], type = "l", lty = 3, col = "black", lwd = 2)
# Add axes
axis(side = 2, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
   labels = c("0", "1", "2", "3", "4"), 
        at = c(0, 1, 2, 3, 4))
axis(side = 4, labels = F, lwd.ticks = 2, 
   at = c(0, 1, 2, 3, 4))
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1, mgp = c(1, 1, 0),
    labels = c(0, 200, 400, 600), at = c(0, 200, 400, 600))
axis(side = 3, labels = F, lwd.ticks = 2, las = 1, cex.axis = 1, 
   at = c(0, 200, 400, 600))
# Add axis labels
mtext('Biomass (OD600)', side = 2, outer = TRUE, cex = 1.5, 
      line = -4, adj = 0.6)
mtext('Time (h)', side = 1, outer = TRUE, cex = 1.5, 
      line = -4, adj = 0.5)
text(400, 3.6, labels = "+Rpf", cex = 1.25)
text(540, 3.6, labels = "-Rpf", cex = 1.25)

# Close Plot Device
dev.off()
graphics.off()
# Show Plot
img <- readPNG("./figures/Fig4.png")
grid.raster(img)
```

## Analyze growth curve parameters

# Load parameters derived from modified Gompertz model

```{r}
g.parms <- read.delim("data/growth.curve/parameters.txt", sep=",", header=TRUE)
```

# Compare summary statistics of growth curve parameters

```{r}
Treatment <- c("Rpf-", "Rpf-","Rpf-","Rpf-", "Rpf+","Rpf+","Rpf+","Rpf+")
g.parms$Treatment <- Treatment

Rpf.plus <- g.parms %>%
  filter(Treatment == "Rpf+")
Rpf.minus <- g.parms %>%
  filter(Treatment == "Rpf-")

# Maximum growth rate (umax)
umax.init <- ddply(g.parms, c("Treatment"), summarise,
                   mean=mean(umax), sem=sd(umax)/sqrt(length(umax)))
umax <- transform(umax.init, lower=mean-sem, upper=mean+sem)

# Lag time
lag.init <- ddply(g.parms, c("Treatment"), summarise,
                   mean=mean(L), sem=sd(L)/sqrt(length(L)))
lag <- transform(lag.init, lower=mean-sem, upper=mean+sem)

# Yield
yield.init <- ddply(g.parms, c("Treatment"), summarise,
                   mean=mean(A), sem=sd(A)/sqrt(length(A)))
yield <- transform(yield.init, lower=mean-sem, upper=mean+sem)

# compare coefficients of variation of growth parameters as function of Rpf
g.parm.CV <- g.parms %>% group_by(Treatment) %>%
   summarise_at(c("L", "umax", "A"), cv)

# Relatively small difference in CV of yield for +Rpf (0.093) vs. -Rpf (0.131)
# CV for umax much higher in -Rpf (0.74) compared to +Rpf (0.0224)
# Similarly, CV for lag is higher in -Rpf (0.114) compared to +Rpf (0.0231)
```

# Statistically test for Rpf effect on growth curve parameters

```{r}
umax.test <- t.test(umax~Treatment, data=g.parms)
# t = 1.4039, df = 3.0013, p-value = 0.2549
lag.test <- t.test(L~Treatment, data=g.parms)
# t = 6.5252, df = 3.0966, p-value = 0.006631
# lag is 1.6 times or 65 % longer without Rpf
# Takes 179 h (or 7.5 days) longer to enter exponential phase
yield.test <- t.test(A~Treatment, data=g.parms)
# t = 1.6802, df = 5.0111, p-value = 0.1536
```


# Statistically test equal variance for growth curve parameter

```{r}
Rpf.plus <- g.parms %>% filter(Treatment == "Rpf+")
Rpf.minus <- g.parms %>% filter(Treatment == "Rpf-")
umax_var_test <- var.test(Rpf.plus$umax, Rpf.minus$umax)
# ratio of variances = 0.000212798, F(3,3) = 0.0002, p = 1.05 × 10−5−5)
lag_var_test <- var.test(Rpf.plus$L, Rpf.minus$L)
# ratio of variances = 0.01609621, F(3,3) = 0.016, p = 0.006737)
yield_var_test <- var.test(Rpf.plus$A, Rpf.minus$A)
# ratio of variances = 0.3848321, F(3,3) = 0.38483, p = 0.4536)
```


## Make figure - growth parameters

```{r}
png(filename="~/GitHub/Rpf/figures/FigS4.png",
    width = 600, height = 1200, res = 96*2) 

plot.new()
par(oma=c(2, 1.5, 0, 1), mar=c(2, 5, 2, 2), mfrow=c(3, 1))

# umax

# Add points = -Rpf
plot(rep(jitter(0.75, amount = 0.1),4), Rpf.minus$umax, 
     ylim = c(-0.03, 0.2), xlim = c(0, 3), pch = 21, col = "lightgrey", 
     bg = "lightgrey", lwd = 2, cex = 1.7, yaxt = "n", xaxt = "n", cex.lab = 2, 
     cex.axis= 1.5, las = 1, ylab = "", xlab = "")
box(lwd=2) 

# Add means
points(0.75, mean(Rpf.minus$umax), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)

# Add error bars
arrows(x0 = 0.75, y0 = Rpf.minus$umax, 
       y1 = LL.95(Rpf.minus$umax), angle = 90, 
       length = 0.1, lwd = 2)
arrows(x0 = 0.75, y0 = Rpf.minus$umax, 
       y1 = UL.95(Rpf.minus$umax), 
       angle = 90, length = 0.1, lwd = 2)

# Add points = +Rpf
points(rep(jitter(2.25, amount = 0.1),4), Rpf.plus$umax, 
       pch = 21, col = "lightgrey", bg = "lightgrey", 
       lwd = 2, cex = 1.7) 

# Add mean = +Rpf
points(2.25, mean(Rpf.plus$umax), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)

# Add error bars
arrows(x0 = 2.25, y0 = Rpf.plus$umax, y1 = LL.95(Rpf.plus$umax), 
       angle = 90, length = 0.1, lwd = 2)
arrows(x0 = 2.25, y0 = Rpf.plus$umax, y1 = UL.95(Rpf.plus$umax), 
       angle = 90, length = 0.1, lwd = 2)

# Add axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("0", "0.1", "0.2"), at = c(0, 0.1, 0.2))
axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at=c(0, 0.1, 0.2), labels = F)
axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(0.75, 2.25), labels = F)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(0.75, 2.25), labels = F)

# Add axis labels
#mtext('-Rpf', side = 1, line = 3, at = 0.75, cex = 1)
#mtext('+Rpf', side = 1, line = 3, at = 2.25, cex = 1)
mtext(expression(mu [max]* ' (hr'^-1*')'), side = 2,
 outer = FALSE, cex = 1.5, line = 4, adj = 0.5)

# yield

# Add points = -Rpf
plot(rep(jitter(0.75, amount = 0.1),4), Rpf.minus$A, 
     ylim = c(2.5, 5.5), xlim = c(0, 3), pch = 21, 
     col = "lightgrey", bg = "lightgrey", lwd = 2, cex = 1.7,
     yaxt = "n", xaxt = "n", cex.lab = 2, 
     cex.axis= 1.5, las = 1, ylab = "", xlab = "")
box(lwd=2) 

# Add means
points(0.75, mean(Rpf.minus$A), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)

# Add error bars
arrows(x0 = 0.75, y0 = Rpf.minus$A, y1 = LL.95(Rpf.minus$A), 
       angle = 90, length = 0.1, lwd = 2)
arrows(x0 = 0.75, y0 = Rpf.minus$A, y1 = UL.95(Rpf.minus$A), 
       angle = 90, length = 0.1, lwd = 2)

# Add points = +Rpf
points(rep(jitter(2.25, amount = 0.1),4), Rpf.plus$A, pch = 21, 
       col = "lightgrey", bg = "lightgrey", lwd = 2, cex = 1.7) 

# Add mean = +Rpf
points(2.25, mean(Rpf.plus$A), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)

# Add error bars
arrows(x0 = 2.25, y0 = Rpf.plus$A, y1 = LL.95(Rpf.plus$A), 
       angle = 90, length = 0.1, lwd = 2)
arrows(x0 = 2.25, y0 = Rpf.plus$A, y1 = UL.95(Rpf.plus$A), 
       angle = 90, length = 0.1, lwd = 2)

# Add axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("3", "4", "5"), at = c(3, 4, 5))
axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at=c(3, 4, 5), labels = F)
axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(0.75, 2.25), labels = F)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(0.75, 2.25), labels = F)

# Add axis labels
#mtext('-Rpf', side = 1, line = 3, at = 0.75, cex = 1)
#mtext('+Rpf', side = 1, line = 3, at = 2.25, cex = 1)
mtext(expression('Yield (OD600)'), side = 2,
      outer = FALSE, cex = 1.5, line = 4, adj = 0.5)

# lag

# Add points = -Rpf
plot(rep(jitter(0.75, amount = 0.1),4), Rpf.minus$L, 
     ylim = c(250, 650), xlim = c(0, 3), pch = 21, 
     col = "lightgrey", bg = "lightgrey", lwd = 2, 
     cex = 1.7, yaxt = "n", xaxt = "n", cex.lab = 2, 
     cex.axis= 1.5, las = 1, ylab = "", xlab = "")
box(lwd=2) 

# Add means
points(0.75, mean(Rpf.minus$L), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)

# Add error bars
arrows(x0 = 0.75, y0 = Rpf.minus$L, y1 = LL.95(Rpf.minus$L), 
       angle = 90, length = 0.1, lwd = 2)
arrows(x0 = 0.75, y0 = Rpf.minus$L, y1 = UL.95(Rpf.minus$L), 
       angle = 90, length = 0.1, lwd = 2)

# Add points = +Rpf
points(rep(jitter(2.25, amount = 0.1),4), Rpf.plus$L, pch = 21, 
       col = "lightgrey", bg = "lightgrey", lwd = 2, cex = 1.7) 

# Add mean = +Rpf
points(2.25, mean(Rpf.plus$L), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)

# Add error bars
arrows(x0 = 2.25, y0 = Rpf.plus$L, y1 = LL.95(Rpf.plus$L), 
       angle = 90, length = 0.1, lwd = 2)
arrows(x0 = 2.25, y0 = Rpf.plus$L, y1 = UL.95(Rpf.plus$L), 
       angle = 90, length = 0.1, lwd = 2)

# Add axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("300", "400", "500", "600"), at = c(300, 400, 500, 600))
axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at=c(300, 400, 500, 600), labels = F)
axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(0.75, 2.25), labels = F)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(0.75, 2.25), labels = F)

# Add axis labels
mtext('-Rpf', side = 1, line = 3, at = 0.75, cex = 1.5)
mtext('+Rpf', side = 1, line = 3, at = 2.25, cex = 1.5)

mtext(expression('Lag time (hr)'), side = 2,
      outer = FALSE, cex = 1.5, line = 4, adj = 0.5)

# Close Plot Device
dev.off()
graphics.off()
# Show Plot
img <- readPNG("./figures/FigS4.png")
grid.raster(img)
```