---
title: "Rpf: generality"
author: "Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

Termination of dormancy in soil bacteria with resuscitation promoting factor (Rpf)

## Clear workspace and set directory

```{r setup}
rm(list=ls())
getwd()
knitr::opts_knit$set(root.dir = '~/../GitHub/Rpf') 
```

## Load packages and functions

```{r}
require("png")
require("dplyr")
require("grid")
require("psych")
```

## Load and process growth data

```{r}
# Load parameters from Gompertz analysis of growth curves
gc.param.1 <- read.delim("data/generality/20160320_generality_gomp_param_output.txt", sep = ",", header = TRUE)
gc.param.2 <- read.delim("data/generality/20160323_generality_gomp_param_output.txt", sep = ",", header = TRUE)

# Load key associated with growth curve experiment 
gc.key <- read.delim("data/generality/generality.gc.well.key.txt", sep = "\t", header = TRUE)

# Merge data and key for data sets
key.1 <- subset(gc.key, Run_Num == "1", select = c(Strain,Gram,Phylum,Genus,Treatment,Rep))
gc.param.1 <- cbind(gc.param.1,key.1)
key.2 <- subset(gc.key, Run_Num == "2", select = c(Strain,Gram,Phylum,Genus,Treatment,Rep))
gc.param.2 <- cbind(gc.param.2,key.2)

# Merge data sets
gc <- rbind(gc.param.1, gc.param.2)

# Remove unused parameters to simplify
gc <- gc[ ,c(-(1:2),-(6:10))]

# Yield parameter (A) sometimes reported as negative; take absolute value
gc$A <- abs(gc$A)

# Time recorded in minutes; report lag and umax in hours
gc$L <-gc$L/60
gc$umax <-gc$umax/60

# Round up small negative estimates of lag for KBS0702 to zero
gc[6,3] = 0
gc[8,3] = 0
```

## Perform two-way ANVOA to test for effect of Rpf on strains

```{r}
# Convert Treatment and Strain from integer/character to factor
gc$Strain <- as.factor(gc$Strain)
gc$Treatment <- as.factor(gc$Treatment)
gc$Rep <- as.factor(gc$Rep)
Rpf <- ifelse(gc$Treatment == "Rpf+", "plus", "minus")
gc <- data.frame(gc, Rpf)
gc$Rpf <- as.factor(gc$Rpf)

# Two-way ANOVA on yield
rpf.aov2.A <- aov(A ~ Treatment + Strain + Treatment*Strain, data = gc)

#                  Df Sum Sq Mean Sq F value   Pr(>F)    
# Treatment         1  0.324  0.3243  41.008 1.36e-08 ***
# Strain           11 12.997  1.1815 149.396  < 2e-16 ***
# Treatment:Strain 11  0.502  0.0456   5.767 1.34e-06 ***
# Residuals        72  0.569  0.0079                     




rpf.aov <- aov(L~ Rpf + Strain + Rpf*Strain, data = gc)
rpf.lm <- lm(L ~ Rpf + Strain + Strain*Rpf, data = gc)
#rpf.lm.umax <- lm(umax ~ Rpf * Strain, data = gc)

#rpf.aov2.umax <- lm(umax ~ Strain, data = gc)
#rpf.aov2.umax < lm(umax ~ Treatment + Strain + Treatment:Strain, data = gc)
#summaryanova<-Anova(rpf.aov2.umax, type = "II")
summary(rpf.aov)
summary(rpf.lm)


#                  Df    Sum Sq   Mean Sq F value  Pr(>F)    
# Treatment         1 5.400e-11 5.350e-11   1.681 0.19888    
# Strain           11 1.274e-08 1.158e-09  36.394 < 2e-16 ***
# Treatment:Strain 11 9.880e-10 8.980e-11   2.822 0.00406 ** 
# Residuals        72 2.291e-09 3.180e-11                  








gc.rpf.plus.strains <- filter(gc, Treatment == "Rpf+") %>%
   select(A, umax, L, Strain) %>%
   group_by(Strain)

gc.rpf.minus.strains <- filter(gc, Treatment == "Rpf-") %>%
   select(A, umax, L, Strain) %>%
   group_by(Strain)

plot(gc.rpf.minus.strains$A,gc.rpf.plus.strains$A)
abline(lm(gc.rpf.plus.strains$A ~ gc.rpf.minus.strains$A), col="red")
abline(0,1)

plot(gc.rpf.minus.strains$umax,gc.rpf.plus.strains$umax)
abline(lm(gc.rpf.plus.strains$umax ~ gc.rpf.minus.strains$umax), col="red")
abline(0,1)

plot(gc.rpf.minus.strains$L,gc.rpf.plus.strains$L)
abline(lm(gc.rpf.plus.strains$L ~ gc.rpf.minus.strains$L), col="red")
abline(0,1)


fold.change.rpf.strains <- data.frame(gc.rpf.plus.strains[,1:3] / 
            gc.rpf.minus.strains[,1:3], gc.rpf.plus.strains[,4]) 

fold.change.rpf.strains <- data.frame((gc.rpf.plus.strains[,1:3] - 
            gc.rpf.minus.strains[,1:3])/gc.rpf.minus.strains[,1:3] , 
            gc.rpf.plus.strains[,4]) 

aov.fold.change.A <- aov(A ~ Strain, data = fold.change.rpf.strains)
# P = 0.174

aov.fold.change.umax <- aov(umax ~ Strain, data = fold.change.rpf.strains)
# P = 0.172

aov.fold.change.L <- aov(L ~ Strain, data = fold.change.rpf.strains)
# P = 0.994

TukeyHSD(aov.fold.change.A)



gc.rpf.plus.strains <- filter(gc, Treatment == "Rpf+") %>%
   select(A, umax, L, Strain) %>%
   group_by(Strain) %>%
   summarize(mean.A = mean(A), mean.umax =mean(umax), mean.L = mean(L)) 

gc.rpf.minus.strains <- filter(gc, Treatment == "Rpf-") %>%
   select(A, umax, L, Strain) %>%
   group_by(Strain) %>%
   summarize(mean.A = mean(A), mean.umax =mean(umax), mean.L = mean(L))

gc.rpf.minus.strains <- filter(gc, Treatment == "Rpf-") %>%
   select(A, umax, L, Strain) %>%
   group_by(Strain) %>%
   summarize(mean.A = mean(A), mean.umax =mean(umax), mean.L = mean(L))




gc.rpf.pool.sd <- select(gc, A, umax, L, Strain) %>%
   group_by(Strain) %>%
   summarize(sd.A = sd(A), sd.umax = sd(umax), sd.L = sd(L)) 

cohen <- (gc.rpf.plus.strains[,2:4] - gc.rpf.minus.strains[,2:4]) / gc.rpf.pool.sd[,2:4]





# Calculate % Biomass yield (A) change
a <- as.data.frame(gc$A[which(gc$Treatment == "Rpf+")])
b <- as.data.frame(gc$A[which(gc$Treatment == "Rpf-")])
df <- c()
df <- as.data.frame(((a-b)/b)*100)

# Calculate % Maximum growth rate (umax) change
c <- as.data.frame(gc$umax[which(gc$Treatment == "Rpf+")])
d <- as.data.frame(gc$umax[which(gc$Treatment == "Rpf-")])
df.u <- as.data.frame(((c-d)/d)*100)
df <- cbind(df, df.u)
# Calculate % Lag time (L) change
e <- gc$L[which(gc$Treatment == "Rpf+")]
f <- gc$L[which(gc$Treatment == "Rpf-")]
df.l <- as.data.frame(((e-f)/f)*100)
df <- cbind(df, df.l)
colnames(df) <- c("A", "umax", "L")

# Add phylogenetic information to dataset
df.info.1 <- as.data.frame(gc[1:24, 4:7])
df.info.2 <- as.data.frame(gc[73:96, 4:7])
df.info <- rbind(df.info.1, df.info.2)
df <- cbind(df, df.info) 


```


















dose.raw <- read.table("data/dose.response/dose.response.txt", header = TRUE, sep="\t")

# Calculate area of plate with bacteria
area <- dose.raw$Pixels / (pi * dose.raw$Plate.Width * dose.raw$Plate.Height)
dose.raw <- data.frame(dose.raw, area)

# Sselect 48 or 96 hrs
biomass <- dose.raw[which(dose.raw$Hrs == 48),]

# Calculate total biomass across plates
sum.area <- sum(biomass$area)

# Calculate relative biomass on each plate with different Rpf concentrations
rel.biomass <- biomass$area/sum.area
dose <- data.frame(biomass, rel.biomass)
```

## Fit growth to Rpf using asymetric Laplace with maxium likelihood

```{r}
A = 0.40 # Maximum biomass (0.4)
X = 40 # Optimum Rpf (45)
K = 0.9 # Kappa; <1 right side decays slowly; >1 left side decays slowly (0.5)
L = -0.04 # Lamdba; peakiness (-0.5); larger = flatter
Z = 0.02 # Error (10)

fit <- mle2(rel.biomass ~ dnorm(mean = a*exp(-(Rpf-x) * l * sign(x - Rpf)* 
            k ^sign(x - Rpf)), sd = z), start = list(a = A, x = X, k = K, l = L, z = Z), 
            data = dose)

#  Coefficients:
#     Estimate Std. Error  z value     Pr(z)    
# a  0.4183468  0.0216615  19.3129 < 2.2e-16 ***
# x 39.0661777         NA       NA        NA    
# k  0.9030598  0.0441071  20.4742 < 2.2e-16 ***
# l -0.0406811  0.0038855 -10.4701 < 2.2e-16 ***
# z  0.0225393  0.0064276   3.5067 0.0004538 ***

# -2 log L: -28.50004 
```

# Plot data and function

```{r}
png(filename="~/GitHub/Rpf/figures/Fig5.png",
    width = 1200, height = 1200, res = 96*2) 

plot.new()
par(mar = c(7, 7, 5, 7))

plot(dose[,2], dose[,8], xlim = c(-25, 185), 
     ylim = c(-0.025, 0.45), type = "p", 
     pch = 22, bg = "grey", col = "black", cex = 2, ylab = "", xlab = "", 
     cex.lab = 1.5, las = 1, lwd = 2, yaxt = "n", xaxt = "n")
box(lwd=2)

# Add ticks and tick labels
axis(side = 2, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
   labels = c(0, 0.1, 0.2, 0.3, 0.4), at = c(0, 0.1, 0.2, 0.3, 0.4))

axis(side = 4, labels = F, lwd.ticks = 2, 
   at = c(0, 0.1, 0.2, 0.3, 0.4))

axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1, mgp = c(3, 1, 0),
    labels = c(0, 50, 100, 150), at = c(0, 50, 100, 150))

axis(side = 3, labels = F, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
   at = c(0, 50, 100, 150))

mtext('Rpf (nM)', side = 1, outer = TRUE, cex = 1.5, 
      line = -4, adj = 0.5)

mtext(expression(paste('Relative biomass')), 
      side = 2, outer = TRUE, cex = 1.5, line = -3, adj = 0.6)

# Plot function
curve(coef(fit)[[1]] * exp(-(x-coef(fit)[[2]]) * coef(fit)[[4]] * sign(coef(fit)[[2]] - x) * 
      coef(fit)[[3]]^sign(coef(fit)[[2]]-x)),from = 0, to = 160, add = TRUE, lty = 2, lwd = 2.5)

# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("./figures/Fig5.png")
grid.raster(img)
```