---
title: "Rpf: dose-response"
author: "Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

Termination of dormancy in soil bacteria with resuscitation promoting factor (Rpf)

## Clear workspace and set directory

```{r setup}
rm(list=ls())
getwd()
setwd("~/GitHub/Rpf")
```

## Load packages and functions

```{r}
require("png")
require("dplyr")
require("grid")
require("plyr")
require("bbmle")
```

## Load and process growth data

Impregnated agar plates with a range of Rpf concentrations and measured areal 
coverage as proxy for growth. 

```{r}
# Load data
dose.raw <- read.table("data/dose.response/dose.response.txt", header = TRUE, sep="\t")

# Calculate area of plate with bacteria
area <- dose.raw$Pixels / (pi * dose.raw$Plate.Width * dose.raw$Plate.Height)
dose.raw <- data.frame(dose.raw, area)

# Sselect 48 or 96 hrs
biomass <- dose.raw[which(dose.raw$Hrs == 48),]

# Calculate total biomass across plates
sum.area <- sum(biomass$area)

# Calculate relative biomass on each plate with different Rpf concentrations
rel.biomass <- biomass$area/sum.area
dose <- data.frame(biomass, rel.biomass)
```

## Fit growth to Rpf using asymetric Laplace with maxium likelihood

```{r}

# The model "fit" includes parameter `a`, which defines maximum growth
# Without `a`, the model can't "reach" the peak

A = 0.40 # Maximum biomass (0.4)
X = 40 # Optimum Rpf (40) or location parameter
K = 0.9 # Kappa (0.9); <1 right side decays slowly; >1 left side decays slowly
L = -0.04 # Lambda (-0.04); larger = peakier; smaller = flatter
Z = 0.1 # Error (0.02)

fit <- mle2(rel.biomass ~ dnorm(mean = a*exp(-(Rpf-x) * l * sign(x - Rpf)* 
            k ^sign(x - Rpf)), sd = z), start = list(a = A, x = X, k = K, l = L, z = Z), 
            data = dose)

#  Coefficients:
#     Estimate Std. Error  z value     Pr(z)    
# a  0.4183468  0.0216615  19.3129 < 2.2e-16 ***
# x 39.0661777         NA       NA        NA    
# k  0.9030598  0.0441071  20.4742 < 2.2e-16 ***
# l -0.0406811  0.0038855 -10.4701 < 2.2e-16 ***
# z  0.0225393  0.0064276   3.5067 0.0004538 ***


# The model "fit2" does not include the parameter `a`
# It doesn't work

M = 37 # Location (40)
K = 1 # Kappa (0.9); shape parameter; 1 = symmetric; <1 = left side steep, >1 right side steeper
L = 0.1 # Lambda (-0.04); larger = peakier; smaller = flatter
Z = 10 # Error (0.02)

fit2 <- mle2(rel.biomass ~ dnorm(mean = (l / k + 1/k) * 
        exp(-(Rpf-m) * l * sign(Rpf-m) * k^sign(Rpf-m)), sd = z),
        start = list(m = M, k = K, l = L, z = Z), data = dose)
```

# Plot data and function

```{r}
png(filename="~/GitHub/Rpf/figures/Fig5.png",
    width = 1200, height = 1200, res = 96*2) 

plot.new()
par(mar = c(7, 7, 5, 7))

plot(dose[,2], dose[,8], xlim = c(-25, 185), 
     ylim = c(-0.025, 0.45), type = "p", 
     pch = 22, bg = "grey", col = "black", cex = 2, ylab = "", xlab = "", 
     cex.lab = 1.5, las = 1, lwd = 2, yaxt = "n", xaxt = "n")
box(lwd=2)

# Add ticks and tick labels
axis(side = 2, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
   labels = c(0, 0.1, 0.2, 0.3, 0.4), at = c(0, 0.1, 0.2, 0.3, 0.4))

axis(side = 4, labels = F, lwd.ticks = 2, 
   at = c(0, 0.1, 0.2, 0.3, 0.4))

axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1, mgp = c(3, 1, 0),
    labels = c(0, 50, 100, 150), at = c(0, 50, 100, 150))

axis(side = 3, labels = F, lwd.ticks = 2, las = 1, cex.axis = 1.25, 
   at = c(0, 50, 100, 150))

mtext('Rpf (nM)', side = 1, outer = TRUE, cex = 1.5, 
      line = -4, adj = 0.5)

mtext(expression(paste('Relative biomass')), 
      side = 2, outer = TRUE, cex = 1.5, line = -3, adj = 0.6)

# Plot function
curve(coef(fit)[[1]] * exp(-(x-coef(fit)[[2]]) * coef(fit)[[4]] * sign(coef(fit)[[2]] - x) * 
      coef(fit)[[3]]^sign(coef(fit)[[2]]-x)),from = 0, to = 160, add = TRUE, lty = 2, lwd = 2.5)

# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("./figures/Fig5.png")
grid.raster(img)
```

