---
title: "Rpf: Mutants"
author: "Jay T. Lennon and Venus Kuo"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

Termination of dormancy in soil bacteria with resuscitation promoting factor (Rpf)

## Clear workspace and set directory

```{r setup}
rm(list=ls())
getwd()
knitr::opts_knit$set(root.dir = '~/../GitHub/Rpf') 
```

## Load packages and functions

```{r}
require("png")
require("dplyr")
require("grid")
require("plyr")

sem <- function(x) sqrt(var(x)/length(x))
LL.95 <- function(x) t.test(x)$conf.int[1]
UL.95 <- function(x) t.test(x)$conf.int[2]
```

## Load growth data

Monitoring dormant KBS0714 populations with different mutations exposed to Rpf.
Mutations included substitution of conserved glutamic acid (E) catalytic site
with other amino acids, namely alanine (A), lysine (K), and gluatimne (Q). 
In addition, we truncated a repeating alanine residue in lectin-like regions, 
which may be involved in Rpf attachment to cell membrane. 

```{r}
mut <- read.delim("data/mutant/mutant.growth.txt", sep="\t", header = TRUE)
```

## Calculate summary statistics

```{r}
mut.mean <- mut %>% group_by(Experiment, Treatment) %>%
 dplyr::summarize(Mean = mean(OD600, na.rm=TRUE))

mut.sem <- mut %>% group_by(Experiment, Treatment) %>%
 dplyr::summarize(SEM = sem(OD600))

mut.LL.95 <- mut %>% group_by(Experiment, Treatment) %>%
  dplyr::summarize(LL.95 = LL.95(OD600))
  
mut.UL.95 <- mut %>% group_by(Experiment, Treatment) %>%
  dplyr::summarize(UL.95 = UL.95(OD600))

mut.table <- data.frame(mut.mean[1], mut.mean[2], mut.mean[3], mut.sem[3],
           mut.LL.95[3], mut.UL.95[3])

colnames(mut.table) <- c("Experiment", "Treatment", "mean", "sem", "LCI", "UCI")
```

# Subset data 
```{r}
mut.E54A.wt <- mut[ which(mut$Experiment == "E54A" & mut$Treatment == "wt"), ]
mut.E54A <- mut[ which(mut$Experiment == "E54A" & mut$Treatment == "mut"), ]

mut.E54K.wt <- mut[ which(mut$Experiment == "E54K" & mut$Treatment == "wt"), ]
mut.E54K <- mut[ which(mut$Experiment == "E54K" & mut$Treatment == "mut"), ]

mut.E54Q.wt <- mut[ which(mut$Experiment == "E54Q" & mut$Treatment == "wt"), ]
mut.E54Q <- mut[ which(mut$Experiment == "E54Q" & mut$Treatment == "mut"), ]

mut.trnc.wt <- mut[ which(mut$Experiment == "trnc" & mut$Treatment == "wt"), ]
mut.trnc <- mut[ which(mut$Experiment == "trnc" & mut$Treatment == "mut"), ]
```

# Statistical test

```{r}
# Mutating glutamic acid (E) at position 54 to alanine (A)
mut.E54A.t.test <- mut[ which(mut$Experiment == "E54A" & mut$Treatment != "control"), ]
t.test.E54A <- t.test(OD600~Treatment, data=mut.E54A.t.test) 
# one-sided t-test: t = -11.576, df = 3.1908, p-value = 0.0005159
# Reduction of biomass = 3.6 fold or 72 % 

# Mutating glutamic acid (E) at position 54 to lysine (K)
mut.E54K.t.test <- mut[ which(mut$Experiment == "E54K" & mut$Treatment != "control"), ]
t.test.E54K <- t.test(OD600~Treatment, data=mut.E54K.t.test) 
# one-sided t-test: t = -9.5292, df = 4.8305, p-value = 0.0002593
# Reduction of biomass = 1.6 fold or 39 % 

# Mutating glutamic acid (E) at position 54 to glutamine (Q)
mut.E54Q.t.test <- mut[ which(mut$Experiment == "E54Q" & mut$Treatment != "control"), ]
t.test.E54Q <- t.test(OD600~Treatment, data=mut.E54Q.t.test) 
# one-sided t-test: t = -5.3561, df = 3.0467, p-value = 0.01224
# Reduction of biomass = 1.3 fold or 25 % 

# Truncating variable linker and LysM Domain (Trnc)
mut.trnc.t.test <- mut[ which(mut$Experiment == "trnc" & mut$Treatment != "control"), ]
t.test.trnc  <- t.test(OD600~Treatment, data=mut.trnc.t.test) 
# one-sided t-test: t = -6.4483, df = 4.29, p-value = 0.002332
# Reduction of biomass = 2.1 fold or 53 % 
```

# Make figure

```{r}
png(filename="~/GitHub/Rpf/figures/Fig4.png",
    width = 1200, height = 1200, res = 96*2) 

plot.new()
par(oma=c(2,4,0,0), mar=c(3,4,4,1), mfrow=c(2,2))

# Panel A: E54A

# Add points
plot(jitter(rep(1.5, length(mut.E54A.wt$OD600)), amount = 0.1), mut.E54A.wt$OD600, 
     ylim = c(-0.2, 3.5), xlim = c(0.5, 5.5), pch = 21, col = "lightgrey", 
     bg = "lightgrey", lwd = 2, cex = 1.7, yaxt = "n", xaxt = "n", cex.lab = 2, 
     cex.axis= 1.5, las = 1, ylab = "", xlab = "")

box(lwd=2) 

points(jitter(rep(4.5, length(mut.E54A$OD600)), amount = 0.1), mut.E54A$OD600, 
       pch = 21, bg = 'lightgrey', col = 'lightgrey', lwd = 2, cex = 1.7)

# Add means
points(1.5, mean(mut.E54A.wt$OD600), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)
points(4.5, mean(mut.E54A$OD600), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)

# Add axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("0.0", "1.0", "2.0", "3.0"), at = c(0, 1, 2, 3))
axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at=c(0, 1, 2, 3), labels = F)
axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1.5, 4.5), labels = F)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1.5, 4.5), labels = F)

# Add error bars
arrows(x0 = c(1.5), y0 = mut.table$mean[which(mut.table$Experiment == "E54A" & 
      mut.table$Treatment == "wt")], y1 = mut.table$LCI[which(mut.table$Experiment 
      == "E54A" & mut.table$Treatment == "wt" )], angle = 90, length = 0.1, lwd = 2)

arrows(x0 = c(1.5), y0 = mut.table$mean[which(mut.table$Experiment == "E54A" & 
      mut.table$Treatment == "wt")], y1 = mut.table$UCI[which(mut.table$Experiment 
      == "E54A" & mut.table$Treatment == "wt" )], angle = 90, length = 0.1, lwd = 2)

arrows(x0 = c(4.5), y0 = mut.table$mean[which(mut.table$Experiment == "E54A" & 
      mut.table$Treatment == "mut")], y1 = mut.table$LCI[which(mut.table$Experiment 
      == "E54A" & mut.table$Treatment == "mut" )], angle = 90, length = 0.1, lwd = 2)

arrows(x0 = c(4.5), y0 = mut.table$mean[which(mut.table$Experiment == "E54A" & 
      mut.table$Treatment == "mut")], y1 = mut.table$UCI[which(mut.table$Experiment 
      == "E54A" & mut.table$Treatment == "mut" )], angle = 90, length = 0.1, lwd = 2)

# Add x-axis labels
mtext('wild', side = 1, line = 1.5, at = 1.5, cex = 1.5)
mtext('type', side = 1, line = 3.0, at = 1.5, cex = 1.5)
mtext('mutant', side = 1, line = 1.5, at = 4.5, cex = 1.5)
mtext('(E54A)', side = 1, line = 3.0, at = 4.5, cex = 1.5)

# Add line corresponding to -Rpf control
E54A.control = mut.table$mean[ which(mut.table$Experiment == "E54A" 
      & mut.table$Treatment == "control") ]
abline(h = E54A.control, col = 'black',  lwd = 2, lty = 2)
text(1, 1.1, labels = expression(~italic("No Rpf")), col = "darkgrey", cex = 1)

# Add panel name and P-value
mtext(expression(~italic("P")~"= 0.0005"), line = -1.5, cex = 1, at = 1.5)
text(5.25, 3.3, labels = "A", cex = 2)


# Panel B: E54K

# Add points
plot(jitter(rep(1.5, length(mut.E54K.wt$OD600)), amount = 0.1), mut.E54K.wt$OD600, 
     ylim = c(-0.2, 3.5), xlim = c(0.5, 5.5), pch = 21, col = 'lightgrey', 
     bg = 'lightgrey', lwd = 2, cex = 1.7, yaxt = "n", xaxt = "n", cex.lab = 2, 
     cex.axis= 1.5, las = 1, ylab = "", xlab = "")

box(lwd=2) 

points(jitter(rep(4.5, length(mut.E54K$OD600)), amount = 0.1), mut.E54K$OD600, 
       pch = 21, bg = 'lightgrey', col = 'lightgrey', lwd = 2, cex = 1.7)

# Add means
points(1.5, mean(mut.E54K.wt$OD600), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)
points(4.5, mean(mut.E54K$OD600), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)

# Add axes
# axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
#      labels = c("0.0", "1.0", "2.0", "3.0"), at = c(0, 1, 2, 3))
axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at=c(0, 1, 2, 3), labels = F)
axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at=c(0, 1, 2, 3), labels = F)
axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1.5, 4.5), labels = F)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1.5, 4.5), labels = F)

# Add error bars
arrows(x0 = c(1.5), y0 = mut.table$mean[which(mut.table$Experiment == "E54K" & 
      mut.table$Treatment == "wt")], y1 = mut.table$LCI[which(mut.table$Experiment 
      == "E54K" & mut.table$Treatment == "wt" )], angle = 90, length = 0.1, lwd = 2)

arrows(x0 = c(1.5), y0 = mut.table$mean[which(mut.table$Experiment == "E54K" & 
      mut.table$Treatment == "wt")], y1 = mut.table$UCI[which(mut.table$Experiment 
      == "E54K" & mut.table$Treatment == "wt" )], angle = 90, length = 0.1, lwd = 2)

arrows(x0 = c(4.5), y0 = mut.table$mean[which(mut.table$Experiment == "E54K" & 
      mut.table$Treatment == "mut")], y1 = mut.table$LCI[which(mut.table$Experiment 
      == "E54K" & mut.table$Treatment == "mut" )], angle = 90, length = 0.1, lwd = 2)

arrows(x0 = c(4.5), y0 = mut.table$mean[which(mut.table$Experiment == "E54K" & 
      mut.table$Treatment == "mut")], y1 = mut.table$UCI[which(mut.table$Experiment 
      == "E54K" & mut.table$Treatment == "mut" )], angle = 90, length = 0.1, lwd = 2)

# Add x-axis labels
mtext('wild', side = 1, line = 1.5, at = 1.5, cex = 1.5)
mtext('type', side = 1, line = 3.0, at = 1.5, cex = 1.5)
mtext('mutant', side = 1, line = 1.5, at = 4.5, cex = 1.5)
mtext('(E54K)', side = 1, line = 3.0, at = 4.5, cex = 1.5)

# Add line corresponding to -Rpf control
E54K.control = mut.table$mean[ which(mut.table$Experiment == "E54K" 
      & mut.table$Treatment == "control") ]
abline(h = E54K.control, col = 'black',  lwd = 2, lty = 2)
text(1, 0.9, labels = expression(~italic("No Rpf")), col = "darkgrey", cex = 1)

# Add panel name and P-value
mtext(expression(~italic("P")~"= 0.0003"), line = -1.5, cex = 1, at = 1.5)
text(5.25, 3.3, labels = "B", cex = 2)


# Panel C: E54K

# Add points
plot(jitter(rep(1.5, length(mut.E54Q.wt$OD600)), amount = 0.1), mut.E54Q.wt$OD600, 
     ylim = c(-0.2, 3.5), xlim = c(0.5, 5.5), pch = 21, col = 'lightgrey', 
     bg = 'lightgrey', lwd = 2, cex = 1.7, yaxt = "n", xaxt = "n", cex.lab = 2, 
     cex.axis= 1.5, las = 1, ylab = "", xlab = "")

box(lwd=2) 

points(jitter(rep(4.5, length(mut.E54Q$OD600)), amount = 0.1), mut.E54Q$OD600, 
       pch = 21, bg = "lightgrey", col = "lightgrey", lwd = 2, cex = 1.7)

# Add means
points(1.5, mean(mut.E54Q.wt$OD600), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)
points(4.5, mean(mut.E54Q$OD600), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)

# Add axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("0.0", "1.0", "2.0", "3.0"), at = c(0, 1, 2, 3))
axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at=c(0, 1, 2, 3), labels = F)
axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1.5, 4.5), labels = F)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1.5, 4.5), labels = F)

# Add error bars
arrows(x0 = c(1.5), y0 = mut.table$mean[which(mut.table$Experiment == "E54Q" & 
      mut.table$Treatment == "wt")], y1 = mut.table$LCI[which(mut.table$Experiment 
      == "E54Q" & mut.table$Treatment == "wt" )], angle = 90, length = 0.1, lwd = 2)

arrows(x0 = c(1.5), y0 = mut.table$mean[which(mut.table$Experiment == "E54Q" & 
      mut.table$Treatment == "wt")], y1 = mut.table$UCI[which(mut.table$Experiment 
      == "E54Q" & mut.table$Treatment == "wt" )], angle = 90, length = 0.1, lwd = 2)

arrows(x0 = c(4.5), y0 = mut.table$mean[which(mut.table$Experiment == "E54Q" & 
      mut.table$Treatment == "mut")], y1 = mut.table$LCI[which(mut.table$Experiment 
      == "E54Q" & mut.table$Treatment == "mut" )], angle = 90, length = 0.1, lwd = 2)

arrows(x0 = c(4.5), y0 = mut.table$mean[which(mut.table$Experiment == "E54Q" & 
      mut.table$Treatment == "mut")], y1 = mut.table$UCI[which(mut.table$Experiment 
      == "E54Q" & mut.table$Treatment == "mut" )], angle = 90, length = 0.1, lwd = 2)

# Add x-axis labels
mtext('wild', side = 1, line = 1.5, at = 1.5, cex = 1.5)
mtext('type', side = 1, line = 3.0, at = 1.5, cex = 1.5)
mtext('mutant', side = 1, line = 1.5, at = 4.5, cex = 1.5)
mtext('(E54Q)', side = 1, line = 3.0, at = 4.5, cex = 1.5)

# Add line corresponding to -Rpf control
E54Q.control = mut.table$mean[ which(mut.table$Experiment == "E54Q" 
      & mut.table$Treatment == "control") ]
abline(h = E54K.control, col = 'black',  lwd = 2, lty = 2)
text(1, 0.9, labels = expression(~italic("No Rpf")), col = "darkgrey", cex = 1)
#legend(0.5, 1.5, expression(italic("No Rpf")), box.col = "white", bg = "white", adj = 0.2)

# Add panel name and P-value
mtext(expression(~italic("P")~"= 0.012"), line = -1.5, cex = 1, at = 1.5)
text(5.25, 3.3, labels = "C", cex = 2)

# Panel D: truncated

# Add points
plot(jitter(rep(1.5, length(mut.trnc.wt$OD600)), amount = 0.1), mut.trnc.wt$OD600, 
     ylim = c(-0.2, 3.5), xlim = c(0.5, 5.5), pch = 21, col = 'lightgrey', 
     bg = 'lightgrey', lwd = 2, cex = 1.7, yaxt = "n", xaxt = "n", cex.lab = 2, 
     cex.axis= 1.5, las = 1, ylab = "", xlab = "")

box(lwd=2) 

points(jitter(rep(4.5, length(mut.trnc$OD600)), amount = 0.1), mut.trnc$OD600, 
       pch = 21, bg = 'lightgrey', col = 'lightgrey', lwd = 2, cex = 1.7)

# Add means
points(1.5, mean(mut.trnc.wt$OD600), pch = 21, col = 'black', 
       bg = "NA", lwd = 2, cex = 2.5)
points(4.5, mean(mut.trnc$OD600), pch = 21, col = 'black', 
       bg = "NA", lwd = 2, cex = 2.5)

# Add axes
# axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
#      labels = c("0.0", "1.0", "2.0", "3.0"), at = c(0, 1, 2, 3))
axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at=c(0, 1, 2, 3), labels = F)
axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at=c(0, 1, 2, 3), labels = F)
axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1.5, 4.5), labels = F)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1.5, 4.5), labels = F)

# Add error bars
arrows(x0 = c(1.5), y0 = mut.table$mean[which(mut.table$Experiment == "trnc" & 
      mut.table$Treatment == "wt")], y1 = mut.table$LCI[which(mut.table$Experiment 
      == "trnc" & mut.table$Treatment == "wt" )], angle = 90, length = 0.1, lwd = 2)

arrows(x0 = c(1.5), y0 = mut.table$mean[which(mut.table$Experiment == "trnc" & 
      mut.table$Treatment == "wt")], y1 = mut.table$UCI[which(mut.table$Experiment 
      == "trnc" & mut.table$Treatment == "wt" )], angle = 90, length = 0.1, lwd = 2)

arrows(x0 = c(4.5), y0 = mut.table$mean[which(mut.table$Experiment == "trnc" & 
      mut.table$Treatment == "mut")], y1 = mut.table$LCI[which(mut.table$Experiment 
      == "trnc" & mut.table$Treatment == "mut" )], angle = 90, length = 0.1, lwd = 2)

arrows(x0 = c(4.5), y0 = mut.table$mean[which(mut.table$Experiment == "trnc" & 
      mut.table$Treatment == "mut")], y1 = mut.table$UCI[which(mut.table$Experiment 
      == "trnc" & mut.table$Treatment == "mut" )], angle = 90, length = 0.1, lwd = 2)

# Add x-axis labels
mtext('wild', side = 1, line = 1.5, at = 1.5, cex = 1.5)
mtext('type', side = 1, line = 3.0, at = 1.5, cex = 1.5)
mtext('mutant', side = 1, line = 1.5, at = 4.5, cex = 1.5)
mtext('(Trnc)', side = 1, line = 3.0, at = 4.5, cex = 1.5)

# Add line corresponding to -Rpf control
trnc.control = mut.table$mean[ which(mut.table$Experiment == "trnc" 
      & mut.table$Treatment == "control") ]
abline(h = trnc.control, col = 'black',  lwd = 2, lty = 2)
text(1, 0.25, labels = expression(~italic("No Rpf")), col = "darkgrey", cex = 1)

# Add panel name and P-value
mtext(expression(~italic("P")~"= 0.002"), line = -1.5, cex = 1, at = 1.5)
text(5.25, 2, labels = "D", cex = 2)

# Add outside Y axis 
mtext(expression('Biomass (OD600)'), side = 2,
      outer = TRUE, cex = 2, line = 1, adj = 0.5)

# Close Plot Device
dev.off()
graphics.off()

img <- readPNG("./figures/Fig4.png")
grid.raster(img)
```